---
title: "Run MED"
author: "Hans Schrieke and Julie Reveillaud"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    lightbox: true
    gallery: false
    fig_caption: true
    highlight: tango
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Paths and libraries setting

```{r message=FALSE, warning=FALSE}
# load main packages, paths and custom functions
source("../../source/main_packages.R")
source("../../source/functions.R")

# load supplementary packages
packages <- c("Biostrings", "seqinr")
invisible(lapply(packages, require, character.only = TRUE))

# create output folders if needed
if (!dir.exists("../../../output/2_Oligotyping")) {dir.create("../../../output/2_Oligotyping")}
if (!dir.exists("../../../output/2_Oligotyping/2A")) {dir.create("../../../output/2_Oligotyping/2A")}
if (!dir.exists("../../../output/2_Oligotyping/2A/Wolbachia")) {dir.create("../../../output/2_Oligotyping/2A/Wolbachia")}
if (!dir.exists("../../../output/2_Oligotyping/2A/Asaia")) {dir.create("../../../output/2_Oligotyping/2A/Asaia")}
if (!dir.exists("../../../output/2_Oligotyping/2A/Chryseobacterium")) {dir.create("../../../output/2_Oligotyping/2A/Chryseobacterium")}
if (!dir.exists("../../../output/2_Oligotyping/2A/Elizabethkingia")) {dir.create("../../../output/2_Oligotyping/2A/Elizabethkingia")}
if (!dir.exists("../../../output/2_Oligotyping/2A/Erwinia")) {dir.create("../../../output/2_Oligotyping/2A/Erwinia")}
if (!dir.exists("../../../output/2_Oligotyping/2A/Legionella")) {dir.create("../../../output/2_Oligotyping/2A/Legionella")}
if (!dir.exists("../../../output/2_Oligotyping/2A/Serratia")) {dir.create("../../../output/2_Oligotyping/2A/Serratia")}

# paths
path_tsv <- "../../../output/1_MED/1B"
path_med <- "../../../output/1_MED/1A/1A_MED_sequences-m0.10-A0-M0-d4"
path_metadata <- "../../../metadata"
path_oligo <- "../../../output/2_Oligotyping/2A"
```

## Create fasta file for each Genus

To run oligotyping, we need to create fasta files for each Genus we want analyze. 

### Wolbachia
```{r warning=FALSE}
# Wolbachia
prepare_data_for_oligotyping(path_RDATA=path_tsv, 
                             path_MED=path_med, 
                             path_NODES=paste0(path_med, "/NODES/"), 
                             path_metadata=path_metadata,
                             metadata="/metadata.csv",
                             genus="Wolbachia", 
                             path_output=paste0(path_oligo, "/wolbachia/")
                             )
```


### Asaia
```{r warning=FALSE}
prepare_data_for_oligotyping(path_RDATA=path_tsv, 
                             path_MED=path_med, 
                             path_NODES=paste0(path_med, "/NODES/"), 
                             path_metadata=path_metadata,
                             metadata="/metadata.csv",
                             genus="Asaia", 
                             path_output=paste0(path_oligo, "/Asaia/")
                             )
```

### Chryseobacterium
```{r warning=FALSE}
prepare_data_for_oligotyping(path_RDATA=path_tsv, 
                             path_MED=path_med, 
                             path_NODES=paste0(path_med, "/NODES/"), 
                             path_metadata=path_metadata,
                             metadata="/metadata.csv",
                             genus="Chryseobacterium", 
                             path_output=paste0(path_oligo, "/Chryseobacterium/")
                             )
```

### Elizabethkingia
```{r warning=FALSE}
prepare_data_for_oligotyping(path_RDATA=path_tsv, 
                             path_MED=path_med, 
                             path_NODES=paste0(path_med, "/NODES/"), 
                             path_metadata=path_metadata,
                             metadata="/metadata.csv",
                             genus="Elizabethkingia", 
                             path_output=paste0(path_oligo, "/Elizabethkingia/")
                             )
```

### Erwinia
```{r warning=FALSE}
prepare_data_for_oligotyping(path_RDATA=path_tsv, 
                             path_MED=path_med, 
                             path_NODES=paste0(path_med, "/NODES/"), 
                             path_metadata=path_metadata,
                             metadata="/metadata.csv",
                             genus="Erwinia", 
                             path_output=paste0(path_oligo, "/Erwinia/")
                             )
```

### Legionella
```{r warning=FALSE}
prepare_data_for_oligotyping(path_RDATA=path_tsv, 
                             path_MED=path_med, 
                             path_NODES=paste0(path_med, "/NODES/"), 
                             path_metadata=path_metadata,
                             metadata="/metadata.csv",
                             genus="Legionella", 
                             path_output=paste0(path_oligo, "/Legionella/")
                             )
```

### Serratia
```{r warning=FALSE}
prepare_data_for_oligotyping(path_RDATA=path_tsv, 
                             path_MED=path_med, 
                             path_NODES=paste0(path_med, "/NODES/"), 
                             path_metadata=path_metadata,
                             metadata="/metadata.csv",
                             genus="Serratia", 
                             path_output=paste0(path_oligo, "/Serratia/")
                             )
```


# Run Oligotyping

To run oligotyping, we used a docker container where oligotyping v2.1 were installed. To install this container, you can follow this great tutorial : https://merenlab.org/2014/09/02/virtualbox/

```{r engine='bash', attr.output='style="max-height: 300px;"'}
# variable for path
OUTPUT="../../../output"
path_data="$OUTPUT/2_Oligotyping/2A"

# do oligotyping for each genus
for genus in Wolbachia Asaia Chryseobacterium Elizabethkingia Legionella Erwinia Serratia

do
  
  # move in the good folder
  if [ $genus == Wolbachia ] 
  then
  cd $path_data/$genus
  else
  cd ../$genus
  fi
  
  # set the number of component used for oligotyping depending on genus
  if [ $genus == Wolbachia ]
  then
  paramC=2
  elif [ $genus == Asaia ]
  then
  paramC=5
    elif [ $genus == Chryseobacterium ]
  then
  paramC=3
    elif [ $genus == Elizabethkingia ]
  then
  paramC=1
    elif [ $genus == Legionella ]
  then
  paramC=4
    elif [ $genus == Erwinia ]
  then
  paramC=8
    elif [ $genus == Serratia ]
  then
  paramC=1
  fi
  
  # run docker session
  docker run -d --name oligotyping --rm -v `pwd`:`pwd` -w `pwd` -it meren/oligotyping:2.1

  # check files 
  docker exec oligotyping ls 

  # get info from fasta 
  docker exec oligotyping o-get-sample-info-from-fasta 2A_oligotyping_${genus}_sequences.fasta

  # fill gaps
  docker exec oligotyping o-pad-with-gaps 2A_oligotyping_${genus}_sequences.fasta

  # get info from fasta to check if all is good after the fill gaps step
  docker exec oligotyping o-get-sample-info-from-fasta 2A_oligotyping_${genus}_sequences.fasta-PADDED-WITH-GAPS

  # Shannon entropy
  docker exec oligotyping entropy-analysis 2A_oligotyping_${genus}_sequences.fasta-PADDED-WITH-GAPS

  # run oligotyping
  docker exec oligotyping oligotype 2A_oligotyping_${genus}_sequences.fasta-PADDED-WITH-GAPS 2A_oligotyping_${genus}_sequences.fasta-PADDED-WITH-GAPS-ENTROPY -c $paramC -M 10

  # kill docker container
  docker container kill oligotyping
  
done
```

# Session info
```{r warning=FALSE, message=FALSE}
sessionInfo()
```