---
title: "Alpha diversity"
author: "Hans Schrieke and Julie Reveillaud"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    lightbox: true
    gallery: false
    fig_caption: true
    highlight: tango
    toc_depth: 6
    code_folding: "show"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Paths and libraries setting

```{r, warning=FALSE, message=FALSE}
# Load main packages, paths and custom functions
source("../../../source/main_packages.R")
source("../../../source/functions.R")

# Load supplementary packages
packages <- c("multcomp", "openxlsx", "kableExtra", "dplyr")
invisible(lapply(packages, require, character.only = TRUE))
```

# Load phyloseq object after decontam
```{r warning=FALSE}
ps.filter <- readRDS("../../../../output/1_MED/1D/1D_MED_phyloseq_decontam.rds")
```

# Alpha diversity plots

## Whole 
```{r warning=FALSE}

# select whole samples
ps.whole <- subset_samples(ps.filter, Organ=="Whole")
ps.whole@sam_data %>% data.frame() -> test

# estimate alpha diversity

measures <- c("Observed", "Chao1", "Shannon")

p1 <- plot_richness(ps.whole, 
                      x="Sample", 
                      color="Strain", 
                      measures=measures, 
                      nrow = 1)

# extract data from richness plot to custom it
df <- p1$data

# change levels and order of species and location for plot
levels(df$Species) <- c("AA", "CP", "CQ")
df$Species <- factor(df$Species, levels = c("AA", "CQ", "CP"))

new_location <- c("Field - Guadeloupe", 
                  "Laboratory - Slab TC (Wolbachia -)",
                  "Laboratory - Lavar",
                  "Field - Bosc", 
                  "Field - Camping Europe")
df$Strain <- factor(df$Strain, levels = new_location)

levels(df$Strain)

# convert df into data.table and set min and max for y-axis
dat <- data.table::data.table(df)
dat[, y_min := value*0.1, by = variable]
dat[, y_max := value*1.1 , by = variable]
dat[dat$variable=="Observed", "y_min"] <- 10
dat[dat$variable=="Chao1", "y_min"] <- 10
dat[dat$variable=="Observed", "y_max"] <- 55
dat[dat$variable=="Chao1", "y_max"] <- 55
#dat[dat$variable=="Shannon", "y_max"] <- 3

# plot
p <- ggplot(dat,aes(Species,value, color=Strain)) +
  facet_wrap(~ variable, drop=T, scale="free_y")+
  geom_boxplot(outlier.colour = NA, alpha=0.8, position = position_dodge2(width=1, preserve="single"))+
  geom_point(size=1, aes(shape=dat$Strain), position=position_dodge(width=0.7, preserve='total'))+
  scale_shape_manual("Strain", values = c(16,17,17,16,16),
                     labels = c("Field - Guadeloupe", 
                                expression(paste("Laboratory - Slab TC (", italic("Wolbachia"), " -)")),
                                "Laboratory - Lavar",
                                "Field - Bosc", 
                                "Field - Camping Europe"))+
  scale_color_manual("Strain", 
                     values=c("#00BF7D", "#5B6BF7", "#00B0F6", "#A3A500", "#F8766D"),
                     labels = c("Field - Guadeloupe", 
                                expression(paste("Laboratory - Slab TC (", italic("Wolbachia"), " -)")),
                                "Laboratory - Lavar",
                                "Field - Bosc", 
                                "Field - Camping Europe"))+
  labs(x="Species", y = "Alpha Diversity Measure")+
  theme(legend.text.align = 0)
p 

# final plot with more space on y-axis
p_bis <- p +
  geom_blank(aes(y = y_min)) +
  geom_blank(aes(y = y_max))
p_bis
```

## Whole with rarefaction
```{r}
# Rarefaction with sample.size=100
ps.whole2 <- ps.whole %>% rarefy_even_depth(rngseed=1, sample.size = 100)
ps.whole2
ps.whole2@sam_data %>% data.frame() -> test2
compute_read_counts(ps.whole2)

p2 <- alpha_div(ps.whole2, measures)
p2


## Rarefaction with sample.size=1000
ps.whole3 <- ps.whole %>% rarefy_even_depth(rngseed=2, sample.size = 1000)
ps.whole3
ps.whole3@sam_data %>% data.frame() -> test3
compute_read_counts(ps.whole3)

p3 <- alpha_div(ps.whole3, measures)
p3

```

## Ovaries
```{r}
# select ovary samples
ps.ovary <- subset_samples(ps.filter, Organ=="Ovary")
sample_data(ps.ovary)$Read_depth <- sample_sums(ps.ovary)

# change factors and order in metadata
metadata.ps <- data.frame(sample_data(ps.ovary))
levels(metadata.ps$Strain)

levels(metadata.ps$Strain) <- c(levels(metadata.ps$Strain), "Laboratory - Lavar")
metadata.ps$Strain[metadata.ps$Strain=="Lavar (labo)"] <- "Laboratory - Lavar"
levels(metadata.ps$Strain)

metadata.ps$Strain <- droplevels(metadata.ps$Strain)
levels(metadata.ps$Strain)

# update metadata in phyloseq object
sample_data(ps.ovary) <- metadata.ps

# estimate alpha diversity
p1 <- plot_richness(ps.ovary, 
                      x="Sample", 
                      color="Strain", 
                      measures=measures, 
                      nrow = 1)

# extract data from richness plot to custom it
df <- p1$data

# changer levels and order of species and location for plot
levels(df$Species) <- c("AA", "CP", "CQ")

df$Species <- factor(df$Species, levels = c("AA", "CQ", "CP"))

new_location <- c("Field - Guadeloupe", 
                  "Laboratory - Slab TC (Wolbachia -)",
                  "Field - Bosc", 
                  "Field - Camping Europe", 
                  "Laboratory - Lavar")
df$Strain <- factor(df$Strain, levels = new_location)

# final plot
p_ovary <- ggplot(df,aes(Species,value,colour=Strain)) +
  facet_wrap(~ variable, drop=T,scale="free")+
  scale_color_manual("Strain", 
                     values=c("#00BF7D", "#A3A500", "#F8766D", "#00B0F6"),
                     labels = c("Field - Guadeloupe", 
                                "Field - Bosc", 
                                "Field - Camping Europe", 
                                "Laboratory - Lavar"))+
  geom_boxplot(outlier.colour = NA, alpha=0.8, position = position_dodge2(width=1, preserve="single")) +
  geom_point(size=2, position=position_dodge(width=0.7, preserve='total'))+
  labs(x="Species", y = "Alpha Diversity Measure")+
    theme(legend.text.align = 0)

p_ovary
```

# Mean of estimated alpha diversity
```{r}
df_score <- data.frame(p$data)

x <- c("Culex pipiens - Field", "Culex pipiens - Bosc", "Culex pipiens - Camping Europe", "Culex pipiens - Lavar (lab)", 
       "Culex quinquefasciatus - Guadeloupe", "Culex quinquefasciatus - Slab TC",
       "Aedes aegypti - Guadeloupe")
y <- c("Observed", "Chao1", "Shannon")

df <- data.frame(matrix(ncol=3, nrow=7))
rownames(df) <- x
colnames(df) <- y

# Aedes aegypti (from Guadeloupe)
df_score[df_score$Species=="AA" & df_score$variable=="Observed", "value"] %>% mean() -> df[7,1]
df_score[df_score$Species=="AA" & df_score$variable=="Chao1", "value"] %>% mean() -> df[7,2]
df_score[df_score$Species=="AA" & df_score$variable=="Shannon", "value"] %>% mean() -> df[7,3]

# Culex pipiens from field
df_score[df_score$Species=="CP" & df_score$variable=="Observed" & df_score$Field=="Field", "value"] %>% mean() -> df[1,1]
df_score[df_score$Species=="CP" & df_score$variable=="Chao1" & df_score$Field=="Field", "value"] %>% mean() -> df[1,2]
df_score[df_score$Species=="CP" & df_score$variable=="Shannon" & df_score$Field=="Field", "value"] %>% mean() -> df[1,3]

# Culex pipiens from labo
df_score[df_score$Species=="CP" & df_score$variable=="Observed" & df_score$Strain=="Laboratory - Lavar", "value"] %>% mean() -> df[4,1]
df_score[df_score$Species=="CP" & df_score$variable=="Chao1" & df_score$Strain=="Laboratory - Lavar", "value"] %>% mean() -> df[4,2]
df_score[df_score$Species=="CP" & df_score$variable=="Shannon" & df_score$Strain=="Laboratory - Lavar", "value"] %>% mean() -> df[4,3]

# Culex pipiens from Bosc
df_score[df_score$Species=="CP" & df_score$variable=="Observed" & df_score$Strain=="Field - Bosc", "value"] %>% mean() -> df[2,1]
df_score[df_score$Species=="CP" & df_score$variable=="Chao1" & df_score$Strain=="Field - Bosc", "value"] %>% mean() -> df[2,2]
df_score[df_score$Species=="CP" & df_score$variable=="Shannon" & df_score$Strain=="Field - Bosc", "value"] %>% mean() -> df[2,3]

# Culex pipiens from Camping Europe
df_score[df_score$Species=="CP" & df_score$variable=="Observed" & df_score$Strain=="Field - Camping Europe", "value"] %>% mean() -> df[3,1]
df_score[df_score$Species=="CP" & df_score$variable=="Chao1" & df_score$Strain=="Field - Camping Europe", "value"] %>% mean() -> df[3,2]
df_score[df_score$Species=="CP" & df_score$variable=="Shannon" & df_score$Strain=="Field - Camping Europe", "value"] %>% mean() -> df[3,3]

# Culex quinquefasciatus from field (Guadeloupe)
df_score[df_score$Species=="CQ" & df_score$variable=="Observed" & df_score$Field=="Field", "value"] %>% mean() -> df[5,1]
df_score[df_score$Species=="CQ" & df_score$variable=="Chao1" & df_score$Field=="Field", "value"] %>% mean() -> df[5,2]
df_score[df_score$Species=="CQ" & df_score$variable=="Shannon" & df_score$Field=="Field", "value"] %>% mean() -> df[5,3]

# Culex quinquefasciatus from labo (Wolbachia-)
df_score[df_score$Species=="CQ" & df_score$variable=="Observed" & df_score$Field=="Lab ", "value"] %>% mean() -> df[6,1]
df_score[df_score$Species=="CQ" & df_score$variable=="Chao1" & df_score$Field=="Lab ", "value"] %>% mean() -> df[6,2]
df_score[df_score$Species=="CQ" & df_score$variable=="Shannon" & df_score$Field=="Lab ", "value"] %>% mean() -> df[6,3]

df %>%
  kbl(booktable=TRUE) %>%
  kable_paper("hover", full_width = F)
```

# Save plots
```{r warning=FALSE}
tiff("../../../../output/1_MED/1E/1Eb_MED_alpha_diversity.tiff", units="in", width=8, height=5, res=300)
p
dev.off()

png("../../../../output/1_MED/1E/1Eb_MED_alpha_diversity.png", units="in", width=8, height=5, res=300)
p
dev.off()

pdf ( "../../../../output/1_MED/1E/1Eb_MED_alpha_diversity.pdf")
p_bis
dev.off()


tiff("../../../../output/1_MED/1E/1Eb_MED_alpha_diversity_rare_100.tiff", units="in", width=8, height=5, res=300)
p2
dev.off()

png("../../../../output/1_MED/1E/1Eb_MED_alpha_diversity_rare_100.png", units="in", width=8, height=5, res=300)
p2
dev.off()

tiff("../../../../output/1_MED/1E/1Eb_MED_alpha_diversity_rare_1000.tiff", units="in", width=8, height=5, res=300)
p3
dev.off()

png("../../../../output/1_MED/1E/1Eb_MED_alpha_diversity_rare_1000.png", units="in", width=8, height=5, res=300)
p3
dev.off()
```


# Statistics 

## Prepare subdatasets for stats
```{r warning=FALSE}
# Whole
df <- p$data
df_culex <- df[df$Species!="AA",]
df_pipiens <- df[df$Species=="CP",]
df_quinque <- df[df$Species=="CQ",]

# Whole rarefied (100)
df2 <- p2$data
df2_culex <- df2[df2$Species!="AA",]
df2_pipiens <- df2[df2$Species=="CP",]
df2_quinque <- df2[df2$Species=="CQ",]

# Whole rarefied (1000)
df3 <- p3$data
df3_culex <- df3[df3$Species!="AA",]
df3_pipiens <- df3[df3$Species=="CP",]
df3_quinque <- df3[df3$Species=="CQ",]

# Wolbachia- vs Wolbachia+
df_culex_wolbachia <- df_culex[df_culex$Species=="CQ" & df_culex$Strain!="Laboratory - Lavar",]

# Wolbachia- vs Wolbachia+ (rarefied 100)
df2_culex_wolbachia <- df2_culex[df_culex$Species=="CQ" & df2_culex$Strain!="Laboratory - Lavar",]

# Wolbachia- vs Wolbachia+ (rarefied 1000)
df3_culex_wolbachia <- df3_culex[df_culex$Species=="CQ" & df3_culex$Strain!="Laboratory - Lavar",]
```

## Influence of depth sequencing

### Effect of Species

#### Observed

```{r warning=FALSE}
# names for final array
names <- c("Group", "Read_depth", "Group:Read_depth", "Residuals")
names2 <- c("Group", "Read_depth", "Residuals")

## Observed
df_observed <- df[df$variable=="Observed"  & df$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_observed)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)

#### Excel
wb <- createWorkbook()
addWorksheet(wb, "Alpha diversity (depth seq)")

writeData(wb, "Alpha diversity (depth seq)", "Effect of species in the field samples", startCol = 1, startRow = 1)
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX", startCol = 1, startRow = 3)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 5)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 6, rowNames = TRUE)

## Observed (rarefied 100)
df_observed_100 <- df2[df$variable=="Observed"  & df2$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_observed_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)

#### Excel
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX (rarefied, 100)", startCol = 1, startRow = 13)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 15)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 16, rowNames = TRUE)

## Observed (rarefied, 1000)
df_observed_1000 <- df3[df3$variable=="Observed"  & df3$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_observed_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)

#### Excel
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX (rarefied, 1000)", startCol = 1, startRow = 23)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 25)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 26, rowNames = TRUE)
```

#### Chao1

```{r warning=FALSE}
# Chao1
df_chao <- df[df$variable=="Chao1"  & df$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_chao)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)

#### Excel
writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX", startCol = 9, startRow = 3)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 5)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 6, rowNames = TRUE)

# Chao1 (rarefied, 100)
df_chao_100 <- df2[df2$variable=="Chao1"  & df2$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_chao_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)

writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX (rarefied, 100)", startCol = 9, startRow = 13)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 15)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 16, rowNames = TRUE)

## Chao (rarefied, 1000)
df_chao_1000 <- df3[df3$variable=="Chao1"  & df3$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_chao_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX (rarefied, 1000)", startCol = 9, startRow = 23)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 25)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 26, rowNames = TRUE)
#saveWorkbook(wb, "Supplementary_Table_2_review_sheet1_30_11_21.xlsx", overwrite = TRUE)
```

#### Shannon
```{r warning=FALSE}
## Shannon
df_shannon <- df[df$variable=="Shannon"  & df$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_shannon)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX", startCol = 17, startRow = 3)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 5)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 6, rowNames = TRUE)

## Shannon (rarefied, 100)
df_shannon_100 <- df2[df2$variable=="Shannon"  & df2$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_shannon_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 100)", startCol = 17, startRow = 13)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 15)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 16, rowNames = TRUE)

## Shannon (rarefied, 1000)
df_shannon_1000 <- df3[df3$variable=="Shannon"  & df3$Field!="Lab ",]

#### Construct linear model
model  <- lm(value~ Species*Read_depth, data = df_shannon_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 1000)", startCol = 17, startRow = 23)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 25)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 26, rowNames = TRUE)

saveWorkbook(wb, "../../../../output/1_MED/1E/Supplementary_Table_2.xlsx", overwrite = TRUE)
```

### Effect of tetracycline

#### Observed
```{r warning=FALSE}
## Observed
df_observed<- df_quinque[df_quinque$variable=="Observed",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_observed)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "Effect of tetracycline in the Culex quinquefasciatus samples", startCol = 1, startRow = 35)
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX", startCol = 1, startRow = 37)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 39)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 40, rowNames = TRUE)

## Observed (rarefied 100)
df_observed_100 <- df2_culex_wolbachia[df2_culex_wolbachia$variable=="Observed",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_observed_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX (rarefied, 100)", startCol = 1, startRow = 47)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 49)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 50, rowNames = TRUE)


## Observed (rarefied, 1000)
df_observed_1000 <- df3_culex_wolbachia[df3_culex_wolbachia$variable=="Observed",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_observed_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX (rarefied, 1000)", startCol = 1, startRow = 57)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 59)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 60, rowNames = TRUE)
```

#### Chao1

```{r warning=FALSE}
## Chao1
df_chao <- df_quinque[df_quinque$variable=="Chao1",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_chao)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX", startCol = 9, startRow = 37)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 39)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 40, rowNames = TRUE)

## Chao1 (rarefied 100)
df_chao_100 <- df2_culex_wolbachia[df2_culex_wolbachia$variable=="Chao1",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_chao_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX (rarefied, 100)", startCol = 9, startRow = 47)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 49)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 50, rowNames = TRUE)


## Chao1 (rarefied, 1000)
df_chao_1000 <- df3_culex_wolbachia[df3_culex_wolbachia$variable=="Chao1",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_chao_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
#ggpubr::ggqqplot(residuals(model))
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "CHAO INDEX (rarefied, 1000)", startCol = 9, startRow = 57)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 59)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 60, rowNames = TRUE)
```

#### Shannon

```{r warning=FALSE}
## Shannon
df_shannon <- df_quinque[df_quinque$variable=="Shannon",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_shannon)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
#ggpubr::ggqqplot(residuals(model))
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX", startCol = 17, startRow = 37)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 39)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 40, rowNames = TRUE)

## Shannon (rarefied 100)
df_shannon_100 <- df2_culex_wolbachia[df2_culex_wolbachia$variable=="Shannon",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_shannon_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 100)", startCol = 17, startRow = 47)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 49)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 50, rowNames = TRUE)

## Shannon (rarefied, 1000)
df_shannon_1000 <- df3_culex_wolbachia[df3_culex_wolbachia$variable=="Shannon",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_shannon_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 1000)", startCol = 17, startRow = 57)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 59)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 60, rowNames = TRUE)

saveWorkbook(wb, "../../../../output/1_MED/1E/Supplementary_Table_2.xlsx", overwrite = TRUE)
```

### Effect of laboratory

#### Observed
```{r warning=FALSE}
## Observed
df_observed <- df_pipiens[df_pipiens$variable=="Observed",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_observed)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "Effect of lab and location in the Culex pipiens samples", startCol = 1, startRow = 68)
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX", startCol = 1, startRow = 70)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 71)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 72, rowNames = TRUE)

## Observed (rarefied 100)
df_observed_100 <- df2_pipiens[df2_pipiens$variable=="Observed",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_observed_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX (rarefied, 100)", startCol = 1, startRow = 79)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 81)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 82, rowNames = TRUE)

## Observed (rarefied, 1000)
df_observed_1000 <- df3_pipiens[df3_pipiens$variable=="Observed", ]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_observed_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX (rarefied, 1000)", startCol = 1, startRow = 89)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 91)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 92, rowNames = TRUE)
```

#### Chao1
```{r warning=FALSE}
## Chao1
df_chao <- df_pipiens[df_pipiens$variable=="Chao1",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_chao)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX", startCol = 9, startRow = 70)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 71)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 72, rowNames = TRUE)

## Chao1 (rarefied 100)
df_chao_100 <- df2_pipiens[df2_pipiens$variable=="Chao1",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_chao_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX (rarefied, 100)", startCol = 9, startRow = 79)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 81)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 82, rowNames = TRUE)

## Chao1 (rarefied, 1000)
df_chao_1000 <- df3_pipiens[df3_pipiens$variable=="Chao1", ]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_chao_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "CHAO INDEX (rarefied, 1000)", startCol = 9, startRow = 89)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 91)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 92, rowNames = TRUE)
```

#### Shannon
```{r warning=FALSE}
## Shannon
df_shannon <- df_pipiens[df_pipiens$variable=="Shannon",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_shannon)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX", startCol = 17, startRow = 70)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 71)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 72, rowNames = TRUE)


## Shannon (rarefied 100)
df_shannon_100 <- df2_pipiens[df2_pipiens$variable=="Shannon",]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_shannon_100)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 100)", startCol = 17, startRow = 79)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 81)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 82, rowNames = TRUE)

## Shannon (rarefied, 1000)
df_shannon_1000 <- df3_pipiens[df3_pipiens$variable=="Shannon", ]

#### Construct linear model
model  <- lm(value~ Strain*Read_depth, data = df_shannon_1000)
anova(model)-> res_anova
res_anova <- res_anova %>% fill_significance("Pr(>F)")

#### Create residuals plot
par(mfrow=c(2,2))
plot(model)


#### Excel
writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 1000)", startCol = 17, startRow = 89)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 91)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 92, rowNames = TRUE)

saveWorkbook(wb, "../../../../output/1_MED/1E/Supplementary_Table_2.xlsx", overwrite = TRUE)
```


## Analysis of alpha diversity

### Effect of species

#### Observed
```{r warning=FALSE}
names <- c("Species", "Residuals")

## Observed
df_observed <- df[df$variable=="Observed"  & df$Field!="Lab ",]
cat("OBSERVED\n\n")

res <- make_anova_tukey(df_observed, df_observed$Species, df_observed$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
addWorksheet(wb, "Alpha diversity")

writeData(wb, "Alpha diversity", "Effect of species in the field samples", startCol = 1, startRow = 1)
writeData(wb, "Alpha diversity", "OBSERVED INDEX", startCol = 1, startRow = 3)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 1, startRow = 5)
writeData(wb, "Alpha diversity", res_anova, startCol = 1, startRow = 6, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 1, startRow = 11)
writeData(wb, "Alpha diversity", res_tukey, startCol = 1, startRow = 12, rowNames = TRUE)
```

#### Chao1
```{r warning=FALSE}                  
## Chao1
df_chao <- df[df$variable=="Chao1"  & df$Field!="Lab ",]
cat("CHAO1\n\n")

res <- make_anova_tukey(df_chao, df_chao$Species, df_chao$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
writeData(wb, "Alpha diversity", "CHAO1 INDEX", startCol = 9, startRow = 3)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 9, startRow = 5)
writeData(wb, "Alpha diversity", res_anova, startCol = 9, startRow = 6, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 9, startRow = 11)
writeData(wb, "Alpha diversity", res_tukey, startCol = 9, startRow = 12, rowNames = TRUE)
```

#### Shannon
```{r warning=FALSE}
## Shannon
df_shannon <- df[df$variable=="Shannon"  & df$Field!="Lab ",]
cat("SHANNON\n\n")
res <- make_anova_tukey(df_shannon, df_shannon$Species, df_shannon$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
writeData(wb, "Alpha diversity", "SHANNON INDEX", startCol = 17, startRow = 3)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 17, startRow = 5)
writeData(wb, "Alpha diversity", res_anova, startCol = 17, startRow = 6, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 17, startRow = 11)
writeData(wb, "Alpha diversity", res_tukey, startCol = 17, startRow = 12, rowNames = TRUE)
```

### Effect of tetracycline
```{r warning=FALSE}
names <- c("Strain", "Residuals")

## Observed
culex_observed <- df_quinque[df_quinque$variable=="Observed",]
cat("OBSERVED\n\n")
res <- make_anova_tukey(culex_observed, culex_observed$Strain, culex_observed$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
writeData(wb, "Alpha diversity", "Effect of tetracycline in the Culex quinquefasciatus samples", startCol = 1, startRow = 20)
writeData(wb, "Alpha diversity", "OBSERVED INDEX", startCol = 1, startRow = 22)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 1, startRow = 24)
writeData(wb, "Alpha diversity", res_anova, startCol = 1, startRow = 25, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 1, startRow = 30)
writeData(wb, "Alpha diversity", res_tukey, startCol = 1, startRow = 31, rowNames = TRUE)
```

#### Chao1
```{r warning=FALSE}
## Chao1
culex_chao <- df_quinque[df_quinque$variable=="Chao1",]
cat("CHAO1\n\n")
res <- make_anova_tukey(culex_chao, culex_chao$Strain, culex_chao$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
writeData(wb, "Alpha diversity", "CHAO1 INDEX", startCol = 9, startRow = 22)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 9, startRow = 24)
writeData(wb, "Alpha diversity", res_anova, startCol = 9, startRow = 25, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 9, startRow = 30)
writeData(wb, "Alpha diversity", res_tukey, startCol = 9, startRow = 31, rowNames = TRUE)
```

#### Shannon

```{r warning=FALSE}
## Shannon
culex_shannon <- df_quinque[df_quinque$variable=="Shannon",]
cat("SHANNON\n\n")
res <- make_anova_tukey(culex_shannon, culex_shannon$Strain, culex_shannon$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
writeData(wb, "Alpha diversity", "SHANNON INDEX", startCol = 17, startRow = 22)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 17, startRow = 24)
writeData(wb, "Alpha diversity", res_anova, startCol = 17, startRow = 25, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 17, startRow = 30)
writeData(wb, "Alpha diversity", res_tukey, startCol = 17, startRow = 31, rowNames = TRUE)
```

### Effect of lab in Culex pipiens samples

#### Observed
```{r warning=FALSE}
## Observed
pipiens_observed <- df_pipiens[df_pipiens$variable=="Observed",]
cat("OBSERVED\n\n")
res <- make_anova_tukey(pipiens_observed, pipiens_observed$Strain, pipiens_observed$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
writeData(wb, "Alpha diversity", "Effect of lab (and location) in the Culex pipiens samples", startCol = 1, startRow = 37)
writeData(wb, "Alpha diversity", "OBSERVED INDEX", startCol = 1, startRow = 39)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 1, startRow = 41)
writeData(wb, "Alpha diversity", res_anova, startCol = 1, startRow = 42, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 1, startRow = 46)
writeData(wb, "Alpha diversity", res_tukey, startCol = 1, startRow = 47, rowNames = TRUE)
```

#### Chao1
```{r warning=FALSE}
## Chao1
pipiens_chao <- df_pipiens[df_pipiens$variable=="Chao1",]
cat("CHAO1\n\n")
res <- make_anova_tukey(pipiens_chao, pipiens_chao$Strain, pipiens_chao$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
writeData(wb, "Alpha diversity", "CHAO1 INDEX", startCol = 9, startRow = 39)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 9, startRow = 41)
writeData(wb, "Alpha diversity", res_anova, startCol = 9, startRow = 42, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 9, startRow = 46)
writeData(wb, "Alpha diversity", res_tukey, startCol = 9, startRow = 47, rowNames = TRUE)
```

#### Shannon
```{r warning=FALSE}
## Shannon
pipiens_shannon <- df_pipiens[df_pipiens$variable=="Shannon",]
cat("SHANNON\n\n")
res <- make_anova_tukey(pipiens_shannon, pipiens_shannon$Strain, pipiens_shannon$value)
res_tukey <- glht.table(res[[2]]) %>% as.data.frame()
res_tukey <- fill_significance(res_tukey, "Pr(>|t|)")
res_anova <- res[[1]] %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

### Excel
writeData(wb, "Alpha diversity", "SHANNON INDEX", startCol = 17, startRow = 39)
writeData(wb, "Alpha diversity", "ANOVA:", startCol = 17, startRow = 41)
writeData(wb, "Alpha diversity", res_anova, startCol = 17, startRow = 42, rowNames = TRUE)
writeData(wb, "Alpha diversity", "Tukey:", startCol = 17, startRow = 46)
writeData(wb, "Alpha diversity", res_tukey, startCol = 17, startRow = 47, rowNames = TRUE)

saveWorkbook(wb, "../../../../output/1_MED/1E/Supplementary_Table_2.xlsx", overwrite = TRUE)
```


# Session info
```{r warning=FALSE, message=FALSE}
sessionInfo()
```


