---
title: "Alpha diversity"
author: "Hans Schrieke and Julie Reveillaud"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    lightbox: true
    gallery: false
    fig_caption: true
    highlight: tango
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Paths and libraries setting

```{r, warning=FALSE, message=FALSE}
# Load main packages, paths and custom functions
source("../../../source/main_packages.R")
source("../../../source/paths.R")
source("../../../source/functions.R")

# Load supplementary packages
packages <- c("multcomp", "openxlsx", "kableExtra")
invisible(lapply(packages, require, character.only = TRUE))
```

# Load phyloseq object after decontam
```{r warning=FALSE}
setwd(path_rdata)
ps.filter <- readRDS("1D_MED_phyloseq_decontam.rds")
```

# Plots

## Whole 
```{r warning=FALSE}

## whole samples
ps.whole <- subset_samples(ps.filter, Organ=="Whole")

## plot

### estimate alpha diversity

measures <- c("Observed", "Chao1", "Shannon")

p1 <- plot_richness(ps.whole, 
                      x="Sample", 
                      color="Strain", 
                      measures=measures, 
                      nrow = 1)

# extract data from richness plot to custom it
df <- p1$data

# changer levels and order of species and location for plot
levels(df$Species) <- c("AA", "CP", "CQ")
df$Species <- factor(df$Species, levels = c("AA", "CQ", "CP"))

new_location <- c("Field - Guadeloupe", 
                  "Laboratory - Slab TC (Wolbachia -)",
                  "Laboratory - Lavar",
                  "Field - Bosc", 
                  "Field - Camping Europe")
df$Strain <- factor(df$Strain, levels = new_location)

levels(df$Strain)
# final plot
p <- ggplot(df,aes(Species,value,colour=Strain)) +
  facet_wrap(~ variable, drop=T,scale="free")+
  scale_color_manual("Strain", 
                     values=c("#00BF7D", "#5B6BF7", "#00B0F6", "#A3A500", "#F8766D"),
                     labels = c("Field - Guadeloupe", 
                                expression(paste("Laboratory - Slab TC (", italic("Wolbachia"), " -)")),
                                "Laboratory - Lavar",
                                "Field - Bosc", 
                                "Field - Camping Europe"))+
  geom_boxplot(outlier.colour = NA, alpha=0.8, position = position_dodge2(width=1, preserve="single")) +
  geom_point(size=2, position=position_dodge(width=0.7, preserve='total'))+
  labs(x="Species", y = "Alpha Diversity Measure")+
    theme(legend.text.align = 0)

p
```

## Whole with rarefaction
```{r}
# Rarefaction with sample.size=100
ps.whole2 <- ps.whole %>% rarefy_even_depth(rngseed=1, sample.size = 100)
ps.whole2
compute_read_counts(ps.whole2)

p2 <- alpha_div3(ps.whole2, measures)
p2


## Rarefaction with sample.size=1000
ps.whole3 <- ps.whole %>% rarefy_even_depth(rngseed=2, sample.size = 1000)
ps.whole3
compute_read_counts(ps.whole3)

p3 <- alpha_div3(ps.whole3, measures)
p3

```

## Ovaries
```{r}
ps.ovary <- subset_samples(ps.filter, Organ=="Ovary")
sample_data(ps.ovary)$Read_depth <- sample_sums(ps.ovary)


## change factors and order in metadata
metadata.ps <- data.frame(sample_data(ps.ovary))
levels(metadata.ps$Strain)

levels(metadata.ps$Strain) <- c(levels(metadata.ps$Strain), "Laboratory - Lavar")
metadata.ps$Strain[metadata.ps$Strain=="Lavar (labo)"] <- "Laboratory - Lavar"
levels(metadata.ps$Strain)

metadata.ps$Strain <- droplevels(metadata.ps$Strain)
levels(metadata.ps$Strain)

## update metadata in phyloseq object
sample_data(ps.ovary) <- metadata.ps


p1 <- plot_richness(ps.ovary, 
                      x="Sample", 
                      color="Strain", 
                      measures=measures, 
                      nrow = 1)

# extract data from richness plot to custom it
df <- p1$data

# changer levels and order of species and location for plot
levels(df$Species) <- c("AA", "CP", "CQ")

df$Species <- factor(df$Species, levels = c("AA", "CQ", "CP"))

new_location <- c("Field - Guadeloupe", 
                  "Laboratory - Slab TC (Wolbachia -)",
                  "Field - Bosc", 
                  "Field - Camping Europe", 
                  "Laboratory - Lavar")
df$Strain <- factor(df$Strain, levels = new_location)

# final plot
p_ovary <- ggplot(df,aes(Species,value,colour=Strain)) +
  facet_wrap(~ variable, drop=T,scale="free")+
  scale_color_manual("Strain", 
                     values=c("#00BF7D", "#A3A500", "#F8766D", "#00B0F6"),
                     labels = c("Field - Guadeloupe", 
                                "Field - Bosc", 
                                "Field - Camping Europe", 
                                "Laboratory - Lavar"))+
  geom_boxplot(outlier.colour = NA, alpha=0.8, position = position_dodge2(width=1, preserve="single")) +
  geom_point(size=2, position=position_dodge(width=0.7, preserve='total'))+
  labs(x="Species", y = "Alpha Diversity Measure")+
    theme(legend.text.align = 0)

p_ovary
```

## Mean of estimated alpha diversity
```{r}
df_score <- data.frame(p$data)

x <- c("Culex pipiens - Field", "Culex pipiens - Bosc", "Culex pipiens - Camping Europe", "Culex pipiens - Lavar (lab)", 
       "Culex quinquefasciatus - Guadeloupe", "Culex quinquefasciatus - Slab TC",
       "Aedes aegypti - Guadeloupe")
y <- c("Observed", "Chao1", "Shannon")

df <- data.frame(matrix(ncol=3, nrow=7))
rownames(df) <- x
colnames(df) <- y

# Aedes aegypti (from Guadeloupe)
df_score[df_score$Species=="AA" & df_score$variable=="Observed", "value"] %>% mean() -> df[7,1]
df_score[df_score$Species=="AA" & df_score$variable=="Chao1", "value"] %>% mean() -> df[7,2]
df_score[df_score$Species=="AA" & df_score$variable=="Shannon", "value"] %>% mean() -> df[7,3]

# Culex pipiens from field
df_score[df_score$Species=="CP" & df_score$variable=="Observed" & df_score$Field=="Field", "value"] %>% mean() -> df[1,1]
df_score[df_score$Species=="CP" & df_score$variable=="Chao1" & df_score$Field=="Field", "value"] %>% mean() -> df[1,2]
df_score[df_score$Species=="CP" & df_score$variable=="Shannon" & df_score$Field=="Field", "value"] %>% mean() -> df[1,3]

# Culex pipiens from labo
df_score[df_score$Species=="CP" & df_score$variable=="Observed" & df_score$Strain=="Laboratory - Lavar", "value"] %>% mean() -> df[4,1]
df_score[df_score$Species=="CP" & df_score$variable=="Chao1" & df_score$Strain=="Laboratory - Lavar", "value"] %>% mean() -> df[4,2]
df_score[df_score$Species=="CP" & df_score$variable=="Shannon" & df_score$Strain=="Laboratory - Lavar", "value"] %>% mean() -> df[4,3]

# Culex pipiens from Bosc
df_score[df_score$Species=="CP" & df_score$variable=="Observed" & df_score$Strain=="Field - Bosc", "value"] %>% mean() -> df[2,1]
df_score[df_score$Species=="CP" & df_score$variable=="Chao1" & df_score$Strain=="Field - Bosc", "value"] %>% mean() -> df[2,2]
df_score[df_score$Species=="CP" & df_score$variable=="Shannon" & df_score$Strain=="Field - Bosc", "value"] %>% mean() -> df[2,3]

# Culex pipiens from Camping Europe
df_score[df_score$Species=="CP" & df_score$variable=="Observed" & df_score$Strain=="Field - Camping Europe", "value"] %>% mean() -> df[3,1]
df_score[df_score$Species=="CP" & df_score$variable=="Chao1" & df_score$Strain=="Field - Camping Europe", "value"] %>% mean() -> df[3,2]
df_score[df_score$Species=="CP" & df_score$variable=="Shannon" & df_score$Strain=="Field - Camping Europe", "value"] %>% mean() -> df[3,3]

# Culex quinquefasciatus from field (Guadeloupe)
df_score[df_score$Species=="CQ" & df_score$variable=="Observed" & df_score$Field=="Field", "value"] %>% mean() -> df[5,1]
df_score[df_score$Species=="CQ" & df_score$variable=="Chao1" & df_score$Field=="Field", "value"] %>% mean() -> df[5,2]
df_score[df_score$Species=="CQ" & df_score$variable=="Shannon" & df_score$Field=="Field", "value"] %>% mean() -> df[5,3]

# Culex quinquefasciatus from labo (Wolbachia-)
df_score[df_score$Species=="CQ" & df_score$variable=="Observed" & df_score$Field=="Lab ", "value"] %>% mean() -> df[6,1]
df_score[df_score$Species=="CQ" & df_score$variable=="Chao1" & df_score$Field=="Lab ", "value"] %>% mean() -> df[6,2]
df_score[df_score$Species=="CQ" & df_score$variable=="Shannon" & df_score$Field=="Lab ", "value"] %>% mean() -> df[6,3]

df %>%
  kbl(booktable=TRUE) %>%
  kable_paper("hover", full_width = F)
```

## Save plots
```{r warning=FALSE}
setwd(path_plot)
tiff("1Eb_MED_alpha_diversity.tiff", units="in", width=8, height=5, res=300)
p
dev.off()

png("1Eb_MED_alpha_diversity.png", units="in", width=8, height=5, res=300)
p
dev.off()
```

# Statistics 

## Effect of depth sequencing (ANOVA / Tukey's test)

### Select species
```{r warning=FALSE}
# Whole
df <- p$data
df_culex <- df[df$Species!="AA",]
df_pipiens <- df[df$Species=="CP",]
df_quinque <- df[df$Species=="CQ",]

# Whole rarefied (100)
df2 <- p2$data
df2_culex <- df2[df2$Species!="AA",]
df2_pipiens <- df2[df2$Species=="CP",]
df2_quinque <- df2[df2$Species=="CQ",]

# Whole rarefied (1000)
df3 <- p3$data
df3_culex <- df3[df3$Species!="AA",]
df3_pipiens <- df3[df3$Species=="CP",]
df3_quinque <- df3[df3$Species=="CQ",]
```


```{r eval=FALSE, include=FALSE}
# Construire le modèle linéaire
model  <- lm(value~ Strain, data = df_observed)
# Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

# Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

# homogénéité de la variance
plot(model, 1)
```

### Effect of species

#### Observed
```{r}
setwd(path_xlsx)

# names for final array
names <- c("Group", "Read_depth", "Group:Read_depth", "Residuals")
names2 <- c("Group", "Read_depth", "Residuals")

## Observed
df_observed <- df[df$variable=="Observed"  & df$Field!="Lab ",]
cat("OBSERVED\n\n")


### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~ Species*Read_depth, data = df_observed)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
shapiro <- rstatix::shapiro_test(residuals(model))
shapiro <- fill_significance(shapiro,"p.value")

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
levene <- df_observed %>% rstatix::levene_test(value ~ Read_depth*Species)
levene <- fill_significance(levene,"p")

df_observed %>% bartlett.test(value~Read_depth, data = df_observed)

### Tab with condition p-vals
x <- c("Normality (Shapiro-Wilk)", "Homogeneity of variance (Levene's test")
y <- c("statistics", "p", "signif")
df_cond <- data.frame(matrix(ncol = 3, nrow = 2))
rownames(df_cond) <- x
colnames(df_cond) <- y

df_cond[1, "statistics"] <- shapiro$statistic
df_cond[2, "statistics"] <- levene$statistic

df_cond[1, "p"] <- shapiro$p.value
df_cond[2, "p"] <- levene$p

df_cond[1, "signif"] <- shapiro$Significance
df_cond[2, "signif"] <- levene$Significance

### Kruskal-Wallis
res.kruskal <- df_observed %>% rstatix::kruskal_test(value ~ Strain)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

df_observed %>% rstatix::kruskal_effsize(value ~ Strain)

### Dunn's test
pwc <- df_observed %>% 
  rstatix::dunn_test(value ~ Strain, p.adjust.method = "bonferroni") 
pwc


#
wb <- createWorkbook()
addWorksheet(wb, "Alpha diversity (depth seq)")

writeData(wb, "Alpha diversity (depth seq)", "Effect of species in the field samples", startCol = 1, startRow = 1)
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX", startCol = 1, startRow = 3)
writeData(wb, "Alpha diversity (depth seq)", "CONDITIONS FOR ANOVA:", startCol = 1, startRow = 5)
writeData(wb, "Alpha diversity (depth seq)", df_cond, startCol = 1, startRow = 6, rowNames = TRUE)
writeData(wb, "Alpha diversity (depth seq)", "Conditions not good for ANOVA", startCol = 1, startRow = 10, rowNames = TRUE)
writeData(wb, "Alpha diversity (depth seq)", "KRUSKAL-WALLIS: ", startCol = 1, startRow = 12)
writeData(wb, "Alpha diversity (depth seq)", res.kruskal, startCol = 1, startRow = 13, rowNames = TRUE)

saveWorkbook(wb, "Supplementary_Table_2_V2.xlsx", overwrite = TRUE)


## Chao1
df_chao <- df[df$variable=="Chao1"  & df$Field!="Lab ",]
cat("CHAO1\n\n")
res <- make_anova(df_chao, df_chao$Species, df_chao$value, df_chao$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX", startCol = 9, startRow = 3)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 5)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 6, rowNames = TRUE)

#saveWorkbook(wb, "Supplementary_Table_2.xlsx", overwrite = TRUE)

## Shannon
df_shannon <- df[df$variable=="Shannon"  & df$Field!="Lab ",]
cat("SHANNON\n\n")
res <- make_anova(df_shannon, df_shannon$Species, df_shannon$value, df_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX", startCol = 17, startRow = 3)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 5)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 6, rowNames = TRUE)

## Shannon (rarefied, 100)
df2_shannon <- df2[df2$variable=="Shannon"  & df2$Field!="Lab ",]
cat("SHANNON\n\n")
res <- make_anova(df2_shannon, df2_shannon$Species, df2_shannon$value, df2_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 100)", startCol = 25, startRow = 3)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 25, startRow = 5)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 25, startRow = 6, rowNames = TRUE)

## Shannon (rarefied, 1000)
df3_shannon <- df3[df3$variable=="Shannon"  & df3$Field!="Lab ",]
cat("SHANNON\n\n")
res <- make_anova(df3_shannon, df3_shannon$Species, df3_shannon$value, df3_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 1000)", startCol = 33, startRow = 3)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 33, startRow = 5)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 33, startRow = 6, rowNames = TRUE)

saveWorkbook(wb, "Supplementary_Table_2_V2.xlsx", overwrite = TRUE)
```

### Effect of tetracycline
```{r warning=FALSE}
setwd(path_xlsx)

# Wolbachia- vs Wolbachia+
df_culex_wolbachia <- df_culex[df_culex$Species=="CQ" & df_culex$Strain!="Laboratory - Lavar",]

# Wolbachia- vs Wolbachia+ (rarefied 100)
df2_culex_wolbachia <- df2_culex[df_culex$Species=="CQ" & df2_culex$Strain!="Laboratory - Lavar",]

# Wolbachia- vs Wolbachia+ (rarefied 1000)
df3_culex_wolbachia <- df3_culex[df_culex$Species=="CQ" & df3_culex$Strain!="Laboratory - Lavar",]

## Observed
culex_observed <- df_culex_wolbachia[df_culex_wolbachia$variable=="Observed",]
cat("OBSERVED\n\n")
res <- make_anova(culex_observed, culex_observed$Strain, culex_observed$value, culex_observed$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "Effect of tetracycline in the Culex quinquefasciatus samples", startCol = 1, startRow = 15)
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX", startCol = 1, startRow = 17)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 19)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 20, rowNames = TRUE)

##saveWorkbook(wb, "Supplementary_Table_2.xlsx", overwrite = TRUE)

## Chao1
culex_chao <- df_culex_wolbachia[df_culex_wolbachia$variable=="Chao1",]
cat("CHAO1\n\n")
res <- make_anova(culex_chao, culex_chao$Strain, culex_chao$value, culex_chao$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX", startCol = 9, startRow = 17)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 19)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 20, rowNames = TRUE)

##saveWorkbook(wb, "Supplementary_Table_2.xlsx", overwrite = TRUE)


## Shannon
culex_shannon <- df_culex_wolbachia[df_culex_wolbachia$variable=="Shannon",]
cat("SHANNON\n\n")
res <- make_anova(culex_shannon, culex_shannon$Strain, culex_shannon$value, culex_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX", startCol = 17, startRow = 17)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 19)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 20, rowNames = TRUE)

## Shannon (rarefied, 100)
culex_shannon <- df2_culex_wolbachia[df2_culex_wolbachia$variable=="Shannon",]
cat("SHANNON\n\n")
res <- make_anova(culex_shannon, culex_shannon$Strain, culex_shannon$value, culex_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 100)", startCol = 25, startRow = 17)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 25, startRow = 19)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 25, startRow = 20, rowNames = TRUE)

## Shannon (rarefied, 1000)
culex_shannon <- df3_culex_wolbachia[df3_culex_wolbachia$variable=="Shannon",]
cat("SHANNON\n\n")
res <- make_anova(culex_shannon, culex_shannon$Strain, culex_shannon$value, culex_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names2
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 1000)", startCol = 33, startRow = 17)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 33, startRow = 19)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 33, startRow = 20, rowNames = TRUE)

##saveWorkbook(wb, "Supplementary_Table_2.xlsx", overwrite = TRUE)
```

### Effect of laboratory
```{r warning=FALSE}
setwd(path_xlsx)

# Culex pipiens

## Observed
pipiens_observed <- df_pipiens[df_pipiens$variable=="Observed",]
cat("OBSERVED\n\n")
res <- make_anova(pipiens_observed, pipiens_observed$Strain, pipiens_observed$value, pipiens_observed$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "Effect of lab and location in the Culex pipiens samples", startCol = 1, startRow = 29)
writeData(wb, "Alpha diversity (depth seq)", "OBSERVED INDEX", startCol = 1, startRow = 31)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 1, startRow = 33)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 1, startRow = 34, rowNames = TRUE)

##saveWorkbook(wb, "Supplementary_Table_2.xlsx", overwrite = TRUE)

## Chao1
pipiens_chao <- df_pipiens[df_pipiens$variable=="Chao1",]
cat("CHAO1\n\n")
res <- make_anova(pipiens_chao, pipiens_chao$Strain, pipiens_chao$value, pipiens_chao$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "CHAO1 INDEX", startCol = 9, startRow = 31)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 9, startRow = 33)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 9, startRow = 34, rowNames = TRUE)

##saveWorkbook(wb, "Supplementary_Table_2.xlsx", overwrite = TRUE)

## Shannon
pipiens_shannon <- df_pipiens[df_pipiens$variable=="Shannon",]
cat("SHANNON\n\n")
res <- make_anova(pipiens_shannon, pipiens_shannon$Strain, pipiens_shannon$value, pipiens_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX", startCol = 17, startRow = 31)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 17, startRow = 33)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 17, startRow = 34, rowNames = TRUE)

## Shannon (rarefied, 100)
pipiens_shannon <- df2_pipiens[df2_pipiens$variable=="Shannon",]
cat("SHANNON\n\n")
res <- make_anova(pipiens_shannon, pipiens_shannon$Strain, pipiens_shannon$value, pipiens_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 100)", startCol = 25, startRow = 31)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 25, startRow = 33)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 25, startRow = 34, rowNames = TRUE)

## Shannon (rarefied, 1000)
pipiens_shannon <- df3_pipiens[df3_pipiens$variable=="Shannon",]
cat("SHANNON\n\n")
res <- make_anova(pipiens_shannon, pipiens_shannon$Strain, pipiens_shannon$value, pipiens_shannon$Read_depth)
res_anova <- res %>% as.data.frame()
res_anova <- fill_significance(res_anova,"Pr(>F)")
rownames(res_anova) <- names
res

writeData(wb, "Alpha diversity (depth seq)", "SHANNON INDEX (rarefied, 1000)", startCol = 33, startRow = 31)
writeData(wb, "Alpha diversity (depth seq)", "ANOVA:", startCol = 33, startRow = 33)
writeData(wb, "Alpha diversity (depth seq)", res_anova, startCol = 33, startRow = 34, rowNames = TRUE)

##saveWorkbook(wb, "Supplementary_Table_2.xlsx", overwrite = TRUE)
```


## True analysis

### Effect of Species

#### Observed
```{r warning=FALSE}
setwd(path_xlsx)

# names for final array
names <- c("Group", "Read_depth", "Group:Read_depth", "Residuals")
names2 <- c("Group", "Read_depth", "Residuals")

## Observed
df_observed <- df[df$variable=="Observed"  & df$Field!="Lab ",]
cat("OBSERVED\n\n")


### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Species*Read_depth, data = df_observed)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
df_observed %>% rstatix::levene_test(value ~ Species)

### Kruskal-Wallis
res.kruskal <- df_observed %>% rstatix::kruskal_test(value ~ Species)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

df_observed %>% rstatix::kruskal_effsize(value ~ Species)

### Dunn's test
pwc <- df_observed %>% 
  rstatix::dunn_test(value ~ Species, p.adjust.method = "bonferroni") 
pwc
```

#### Chao1

```{r}
## Chao1
df_chao <- df[df$variable=="Chao1"  & df$Field!="Lab ",]
cat("OBSERVED\n\n")


### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Species*Read_depth, data = df_chao)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
df_chao %>% rstatix::levene_test(value ~ Species)

### Kruskal-Wallis
res.kruskal <- df_chao %>% rstatix::kruskal_test(value ~ Species)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

df_chao %>% rstatix::kruskal_effsize(value ~ Species)

### Dunn's test
pwc <- df_chao %>% 
  rstatix::dunn_test(value ~ Species, p.adjust.method = "bonferroni") 
pwc
```

#### Shannon

```{r}
## Shannon
df_shannon <- df[df$variable=="Shannon"  & df$Field!="Lab ",]
cat("OBSERVED\n\n")

### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Species*Read_depth, data = df_shannon)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
df_shannon %>% rstatix::levene_test(value ~ Species)

### Kruskal-Wallis
res.kruskal <- df_shannon %>% rstatix::kruskal_test(value ~ Species)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

df_shannon %>% rstatix::kruskal_effsize(value ~ Species)

### Dunn's test
pwc <- df_shannon %>% 
  rstatix::dunn_test(value ~ Species, p.adjust.method = "bonferroni") 
pwc
```

### Effect of tetracycline

```{r}
setwd(path_xlsx)

# Wolbachia- vs Wolbachia+
df_culex_wolbachia <- df_culex[df_culex$Species=="CQ" & df_culex$Strain!="Laboratory - Lavar",]

# Wolbachia- vs Wolbachia+ (rarefied 100)
df2_culex_wolbachia <- df2_culex[df_culex$Species=="CQ" & df2_culex$Strain!="Laboratory - Lavar",]

# Wolbachia- vs Wolbachia+ (rarefied 1000)
df3_culex_wolbachia <- df3_culex[df_culex$Species=="CQ" & df3_culex$Strain!="Laboratory - Lavar",]
```

#### Observed
```{r}
## Observed
culex_observed <- df_culex_wolbachia[df_culex_wolbachia$variable=="Observed",]
cat("OBSERVED\n\n")

### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Strain, data = culex_observed)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
culex_observed %>% rstatix::levene_test(value ~ Strain)

### Kruskal-Wallis
res.kruskal <- culex_observed %>% rstatix::kruskal_test(value ~ Strain)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

culex_observed %>% rstatix::kruskal_effsize(value ~ Strain)

### Dunn's test
pwc <- culex_observed %>% 
  rstatix::dunn_test(value ~ Strain, p.adjust.method = "bonferroni") 
pwc
```

#### Chao1
```{r}
## Chao1
culex_chao <- df_culex_wolbachia[df_culex_wolbachia$variable=="Chao1",]
cat("OBSERVED\n\n")

### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Strain, data = culex_chao)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
culex_chao %>% rstatix::levene_test(value ~ Strain)

### Kruskal-Wallis
res.kruskal <- culex_chao %>% rstatix::kruskal_test(value ~ Strain)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

culex_chao %>% rstatix::kruskal_effsize(value ~ Strain)

### Dunn's test
pwc <- culex_chao %>% 
  rstatix::dunn_test(value ~ Strain, p.adjust.method = "bonferroni") 
pwc
```
#### Shannon
```{r}
## Shannon
culex_shannon <- df_culex_wolbachia[df_culex_wolbachia$variable=="Shannon",]
cat("OBSERVED\n\n")

### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Strain, data = culex_shannon)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
culex_shannon %>% rstatix::levene_test(value ~ Strain)

### Kruskal-Wallis
res.kruskal <- culex_shannon %>% rstatix::kruskal_test(value ~ Strain)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

culex_shannon %>% rstatix::kruskal_effsize(value ~ Strain)

### Dunn's test
pwc <- culex_shannon %>% 
  rstatix::dunn_test(value ~ Strain, p.adjust.method = "bonferroni") 
pwc
```

### Effect of laboratory

#### Observed
```{r}
pipiens_observed <- df_pipiens[df_pipiens$variable=="Observed",]
cat("OBSERVED\n\n")

### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Strain, data = pipiens_observed)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
pipiens_observed %>% rstatix::levene_test(value ~ Strain)

### Kruskal-Wallis
res.kruskal <- pipiens_observed %>% rstatix::kruskal_test(value ~ Strain)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

pipiens_observed %>% rstatix::kruskal_effsize(value ~ Strain)

### Dunn's test
pwc <- pipiens_observed %>% 
  rstatix::dunn_test(value ~ Strain, p.adjust.method = "bonferroni") 
pwc

```

#### Chao1
```{r}
pipiens_chao <- df_pipiens[df_pipiens$variable=="Chao1",]
cat("OBSERVED\n\n")

### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Strain, data = pipiens_chao)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
pipiens_chao %>% rstatix::levene_test(value ~ Strain)

### Kruskal-Wallis
res.kruskal <- pipiens_chao %>% rstatix::kruskal_test(value ~ Strain)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

pipiens_chao %>% rstatix::kruskal_effsize(value ~ Strain)

### Dunn's test
pwc <- pipiens_chao %>% 
  rstatix::dunn_test(value ~ Strain, p.adjust.method = "bonferroni") 
pwc

```

#### Shannon
```{r}
pipiens_shannon <- df_pipiens[df_pipiens$variable=="Shannon",]
cat("OBSERVED\n\n")

### Conditions pour ANOVA

#### Construire le modèle linéaire
model  <- lm(value~Strain, data = pipiens_shannon)

#### Créer un QQ plot des résidus
ggpubr::ggqqplot(residuals(model))

#### Calculer le test de normalité de Shapiro-Wilk
rstatix::shapiro_test(residuals(model))

#### Homogénéité de la variance
plot(model, 1)

#### Test de Levene
pipiens_shannon %>% rstatix::levene_test(value ~ Strain)

### Kruskal-Wallis
res.kruskal <- pipiens_shannon %>% rstatix::kruskal_test(value ~ Strain)
res.kruskal <- fill_significance(res.kruskal,"p")
res.kruskal

pipiens_shannon %>% rstatix::kruskal_effsize(value ~ Strain)

### Dunn's test
pwc <- pipiens_shannon %>% 
  rstatix::dunn_test(value ~ Strain, p.adjust.method = "bonferroni") 
pwc

```

