---
title: "Remove contaminants and filter samples with less than 1000 reads"
author: "Hans Schrieke and Julie Reveillaud"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    lightbox: true
    gallery: false
    fig_caption: true
    highlight: tango
    toc_depth: 3
    code_folding: "show"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Paths and libraries setting

```{r, warning=FALSE, message=FALSE}
# Load main packages, paths and custom functions
source("../../source/main_packages.R")
source("../../source/functions.R")

# Load supplementary packages
packages <- c("decontam", "kableExtra", "microbiome", "cowplot")
invisible(lapply(packages, require, character.only = TRUE))

# create output folders if needed
if (!dir.exists("../../../output/1_MED/1D")) {dir.create("../../../output/1_MED/1D")}
```

# Import MED phyloseq object
```{r warning=FALSE}
ps <- readRDS("../../../output/1_MED/1C/1C_MED_phyloseq.rds")
```

# Run Decontam 

## Detect contaminants
```{r warning=FALSE}
# set threshold
threshold=0.6
  
# Preprocess
df <- as.data.frame(sample_data(ps)) # Put sample_data into a ggplot-friendly data.frame
df <- data.frame(sample_data(ps)) # Put sample_data into a ggplot-friendly data.frame
df <- df[order(df$Read_depth),]
df$Index <- seq(nrow(df))

# read depth 
p <- ggplot(data=df, aes(x=Index, y=Read_depth, color=Control)) + geom_point()
p 

# Identify contaminants - Prevalence method
sample_data(ps)$is.neg <- sample_data(ps)$Control == "Control sample"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg", threshold=threshold)

# table of positive and false contaminant
table(contamdf.prev$contaminant)

# contaminant nodes
decontam_asv_MED  <- row.names(contamdf.prev[contamdf.prev$contaminant == TRUE, ])

# make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Control == "Control sample", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Control == "True sample", ps.pa)

# make dataframe from phyloseq objects of presence-absence
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                    contaminant=contamdf.prev$contaminant)

# plot
p2 <- ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")

p2

# write the contaminant ASV table detected with prevalence
tax <- tax_table(ps) %>% as.matrix()
decontam_df_MED <- tax[row.names(tax) %in% decontam_asv_MED , ]

# print contaminants
decontam_df_MED   %>% 
  kbl() %>%
  kable_paper("hover", full_width = F)

# histogram of the scores
p <- hist(contamdf.prev$p, 100, ylim = c(0,25), xlim = c(0,1), main="", xlab="p", ylab="Frequency")
```

## Save contaminant table
```{r eval=FALSE, warning=FALSE, include=FALSE}
write.table(contam, "../../../output/1_MED/1D/1D_decontam_contaminants.tsv", sep="\t", quote=F, col.names=NA)
```

## Remove contaminants from the phyloseq object
```{r}
# remove contaminants ASV
alltaxa <- taxa_names(ps)
decontam_taxa <- alltaxa[!(alltaxa %in% decontam_asv_MED)]
ps2 <- prune_taxa(decontam_taxa, ps)

# check ps 
ps2 <- check_ps(ps2)
ps2
compute_read_counts(ps2)
compute_read_counts(ps)-compute_read_counts(ps2)

ps <- ps2
```

## Remove blanks 
```{r warning=FALSE}
# blanks
ps.blanks <- subset_samples(ps, Strain=="Blank")
ps.blanks <- check_ps(ps.blanks)
ps.blanks
compute_read_counts(ps.blanks)

# supprimer blanks
ps.filter <- subset_samples(ps, Strain!="Blank")
ps.filter <- check_ps(ps.filter)
ps.filter
compute_read_counts(ps.filter)

# check nreads
summarize_phyloseq(ps.filter)
```

# Counts

## Create dataframe

```{r}
x <- c("Culex pipiens", "Field - Bosc", "Field - Camping Europe", "Laboratory - Lavar", 
       "Culex quinquefasciatus", "Field - Guadeloupe", "Laboratory - Slab TC (Wolbachia -)",
       "Aedes aegypti (Guadeloupe)",
       "Total")
y <- c("Reads", "Oligotypes", "Samples")

df <- data.frame(matrix(ncol=3, nrow=9))
rownames(df) <- x
colnames(df) <- y

df2 <- df
df3 <- df
```


## Whole + ovary

```{r}
# Culex pipiens
ps.pipiens <- subset_samples(ps.filter, Species=="Culex pipiens")
ps.pipiens <- check_ps(ps.pipiens)

## All Strain
### oligotype
nrow(ps.pipiens@otu_table) -> df["Culex pipiens", "Oligotypes"]
### samples
nrow(ps.pipiens@sam_data) -> df["Culex pipiens", "Samples"]
### reads
compute_read_counts(ps.pipiens) -> df["Culex pipiens", "Reads"]


## Bosc
### oligotype
nrow(otu_table(ps.pipiens %>% 
                 subset_samples(Strain=="Field - Bosc") %>% 
                 check_ps())) -> df["Field - Bosc", "Oligotypes"]
### samples
nrow(ps.pipiens@sam_data %>% subset_samples(Strain=="Field - Bosc")) -> df["Field - Bosc", "Samples"]
### reads
compute_read_counts(ps.pipiens %>% subset_samples(Strain=="Field - Bosc")) -> df["Field - Bosc", "Reads"]


## Camping Europe
### oligotype
nrow(otu_table(ps.pipiens %>% 
                 subset_samples(Strain=="Field - Camping Europe") %>% 
                 check_ps())) -> df["Field - Camping Europe", "Oligotypes"]
### samples
nrow(ps.pipiens@sam_data %>% subset_samples(Strain=="Field - Camping Europe")) -> df["Field - Camping Europe", "Samples"]
### reads
compute_read_counts(ps.pipiens %>% subset_samples(Strain=="Field - Camping Europe")) -> df["Field - Camping Europe", "Reads"]

## Lavar (labo)
### oligotype
nrow(otu_table(ps.pipiens %>% 
                 subset_samples(Strain=="Laboratory - Lavar") %>% 
                 check_ps())) -> df["Laboratory - Lavar", "Oligotypes"]
### samples
nrow(ps.pipiens@sam_data %>% subset_samples(Strain=="Laboratory - Lavar")) -> df["Laboratory - Lavar", "Samples"]
### reads
compute_read_counts(ps.pipiens %>% subset_samples(Strain=="Laboratory - Lavar")) -> df["Laboratory - Lavar", "Reads"]


# Culex quinquefasciatus
ps.quinque <- subset_samples(ps.filter, Species=="Culex quinquefasciatus")
ps.quinque <- check_ps(ps.quinque)

## All Strain
### oligotype
nrow(ps.quinque@otu_table) -> df["Culex quinquefasciatus", "Oligotypes"]
### samples
nrow(ps.quinque@sam_data) -> df["Culex quinquefasciatus", "Samples"]
### reads
compute_read_counts(ps.quinque) -> df["Culex quinquefasciatus", "Reads"]


## Guadeloupe
### oligotype
nrow(otu_table(ps.quinque %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df["Field - Guadeloupe", "Oligotypes"]
### samples
nrow(ps.quinque@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df["Field - Guadeloupe", "Samples"]
### reads
compute_read_counts(ps.quinque %>% subset_samples(Strain=="Field - Guadeloupe")) -> df["Field - Guadeloupe", "Reads"]


## Wolbachia -
### oligotype
nrow(otu_table(ps.quinque %>% 
                 subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)") %>% 
                 check_ps())) -> df["Laboratory - Slab TC (Wolbachia -)", "Oligotypes"]
### samples
nrow(ps.quinque@sam_data %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df["Laboratory - Slab TC (Wolbachia -)", "Samples"]
### reads
compute_read_counts(ps.quinque %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df["Laboratory - Slab TC (Wolbachia -)", "Reads"]


# Aedes aegypti
ps.aedes <- subset_samples(ps.filter, Species=="Aedes aegypti")
ps.aedes <- check_ps(ps.aedes)

## Guadeloupe
### oligotype
nrow(otu_table(ps.aedes %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df["Aedes aegypti (Guadeloupe)", "Oligotypes"]
### samples
nrow(ps.aedes@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df["Aedes aegypti (Guadeloupe)", "Samples"]
### reads
compute_read_counts(ps.aedes %>% subset_samples(Strain=="Field - Guadeloupe")) -> df["Aedes aegypti (Guadeloupe)", "Reads"]

# Total
### oligotype
nrow(ps.filter@otu_table) -> df["Total", "Oligotypes"]
### samples
nrow(ps.filter@sam_data) -> df["Total", "Samples"]
### reads
compute_read_counts(ps.filter) -> df["Total", "Reads"]

df %>% 
  kbl() %>%
  kable_paper("hover", full_width = F)
```

## Whole

```{r}
# Culex pipiens
ps.pipiens.whole <- ps.pipiens %>% subset_samples(Organ=="Whole")
ps.pipiens.whole <- ps.pipiens.whole %>% check_ps()

## All Strain
### oligotype
nrow(ps.pipiens.whole@otu_table) -> df2["Culex pipiens", "Oligotypes"]
### samples
nrow(ps.pipiens.whole@sam_data) -> df2["Culex pipiens", "Samples"]
### reads
compute_read_counts(ps.pipiens.whole) -> df2["Culex pipiens", "Reads"]


## Bosc
### oligotype
nrow(otu_table(ps.pipiens.whole %>% 
                 subset_samples(Strain=="Field - Bosc") %>% 
                 check_ps())) -> df2["Field - Bosc", "Oligotypes"]
### samples
nrow(ps.pipiens.whole@sam_data %>% subset_samples(Strain=="Field - Bosc")) -> df2["Field - Bosc", "Samples"]
### reads
compute_read_counts(ps.pipiens.whole %>% subset_samples(Strain=="Field - Bosc")) -> df2["Field - Bosc", "Reads"]


## Camping Europe
### oligotype
nrow(otu_table(ps.pipiens.whole %>% 
                 subset_samples(Strain=="Field - Camping Europe") %>% 
                 check_ps())) -> df2["Field - Camping Europe", "Oligotypes"]
### samples
nrow(ps.pipiens.whole@sam_data %>% subset_samples(Strain=="Field - Camping Europe")) -> df2["Field - Camping Europe", "Samples"]
### reads
compute_read_counts(ps.pipiens.whole %>% subset_samples(Strain=="Field - Camping Europe")) -> df2["Field - Camping Europe", "Reads"]

## Lavar (labo)
### oligotype
nrow(otu_table(ps.pipiens.whole %>% 
                 subset_samples(Strain=="Laboratory - Lavar") %>% 
                 check_ps())) -> df2["Laboratory - Lavar", "Oligotypes"]
### samples
nrow(ps.pipiens.whole@sam_data %>% subset_samples(Strain=="Laboratory - Lavar")) -> df2["Laboratory - Lavar", "Samples"]
### reads
compute_read_counts(ps.pipiens.whole %>% subset_samples(Strain=="Laboratory - Lavar")) -> df2["Laboratory - Lavar", "Reads"]



# Culex quinquefasciatus
ps.quinque.whole <- subset_samples(ps.quinque, Organ=="Whole")
ps.quinque.whole <- check_ps(ps.quinque.whole)

## All Strain
### oligotype
nrow(ps.quinque.whole@otu_table) -> df2["Culex quinquefasciatus", "Oligotypes"]
### samples
nrow(ps.quinque.whole@sam_data) -> df2["Culex quinquefasciatus", "Samples"]
### reads
compute_read_counts(ps.quinque.whole) -> df2["Culex quinquefasciatus", "Reads"]


## Guadeloupe
### oligotype
nrow(otu_table(ps.quinque.whole %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df2["Field - Guadeloupe", "Oligotypes"]
### samples
nrow(ps.quinque.whole@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df2["Field - Guadeloupe", "Samples"]
### reads
compute_read_counts(ps.quinque.whole %>% subset_samples(Strain=="Field - Guadeloupe")) -> df2["Field - Guadeloupe", "Reads"]


## Wolbachia -
### oligotype
nrow(otu_table(ps.quinque.whole %>% 
                 subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)") %>% 
                 check_ps())) -> df2["Laboratory - Slab TC (Wolbachia -)", "Oligotypes"]
### samples
nrow(ps.quinque.whole@sam_data %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df2["Laboratory - Slab TC (Wolbachia -)", "Samples"]
### reads
compute_read_counts(ps.quinque.whole %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df2["Laboratory - Slab TC (Wolbachia -)", "Reads"]



# Aedes aegypti
ps.aedes.whole <- subset_samples(ps.aedes, Organ=="Whole")
ps.aedes.whole <- check_ps(ps.aedes.whole)

## Guadeloupe
### oligotype
nrow(otu_table(ps.aedes.whole %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df2["Aedes aegypti (Guadeloupe)", "Oligotypes"]
### samples
nrow(ps.aedes.whole@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df2["Aedes aegypti (Guadeloupe)", "Samples"]
### reads
compute_read_counts(ps.aedes.whole %>% subset_samples(Strain=="Field - Guadeloupe")) -> df2["Aedes aegypti (Guadeloupe)", "Reads"]


# Total
ps.whole <- ps.filter %>% subset_samples(Organ=="Whole")
ps.whole <- ps.whole %>% check_ps()
### oligotype
nrow(ps.whole@otu_table) -> df2["Total", "Oligotypes"]
### samples
nrow(ps.whole@sam_data) -> df2["Total", "Samples"]
### reads
compute_read_counts(ps.whole) -> df2["Total", "Reads"]

df2 %>% 
  kbl() %>%
  kable_paper("hover", full_width = F)
```

## Ovary 

```{r}
# Culex pipiens
ps.pipiens.ovary <- ps.pipiens %>% subset_samples(Organ=="Ovary")
ps.pipiens.ovary <- ps.pipiens.ovary %>% check_ps()

## All Strain
### oligotype
nrow(ps.pipiens.ovary@otu_table) -> df3["Culex pipiens", "Oligotypes"]
### samples
nrow(ps.pipiens.ovary@sam_data) -> df3["Culex pipiens", "Samples"]
### reads
compute_read_counts(ps.pipiens.ovary) -> df3["Culex pipiens", "Reads"]


## Bosc
### oligotype
nrow(otu_table(ps.pipiens.ovary %>% 
                 subset_samples(Strain=="Field - Bosc") %>% 
                 check_ps())) -> df3["Field - Bosc", "Oligotypes"]
### samples
nrow(ps.pipiens.ovary@sam_data %>% subset_samples(Strain=="Field - Bosc")) -> df3["Field - Bosc", "Samples"]
### reads
compute_read_counts(ps.pipiens.ovary %>% subset_samples(Strain=="Field - Bosc")) -> df3["Field - Bosc", "Reads"]


## Camping Europe
### oligotype
nrow(otu_table(ps.pipiens.ovary %>% 
                 subset_samples(Strain=="Field - Camping Europe") %>% 
                 check_ps())) -> df3["Field - Camping Europe", "Oligotypes"]
### samples
nrow(ps.pipiens.ovary@sam_data %>% subset_samples(Strain=="Field - Camping Europe")) -> df3["Field - Camping Europe", "Samples"]
### reads
compute_read_counts(ps.pipiens.ovary %>% subset_samples(Strain=="Field - Camping Europe")) -> df3["Field - Camping Europe", "Reads"]

## Lavar (labo)
### oligotype
nrow(otu_table(ps.pipiens.ovary %>% 
                 subset_samples(Strain=="Laboratory - Lavar") %>% 
                 check_ps())) -> df3["Laboratory - Lavar", "Oligotypes"]
### samples
nrow(ps.pipiens.ovary@sam_data %>% subset_samples(Strain=="Laboratory - Lavar")) -> df3["Laboratory - Lavar", "Samples"]
### reads
compute_read_counts(ps.pipiens.ovary %>% subset_samples(Strain=="Laboratory - Lavar")) -> df3["Laboratory - Lavar", "Reads"]



# Culex quinquefasciatus
ps.quinque.ovary <- subset_samples(ps.quinque, Organ=="Ovary")
ps.quinque.ovary <- check_ps(ps.quinque.ovary)

## All Strain
### oligotype
nrow(ps.quinque.ovary@otu_table) -> df3["Culex quinquefasciatus", "Oligotypes"]
### samples
nrow(ps.quinque.ovary@sam_data) -> df3["Culex quinquefasciatus", "Samples"]
### reads
compute_read_counts(ps.quinque.ovary) -> df3["Culex quinquefasciatus", "Reads"]


## Guadeloupe
### oligotype
nrow(otu_table(ps.quinque.ovary %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df3["Field - Guadeloupe", "Oligotypes"]
### samples
nrow(ps.quinque.ovary@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df3["Field - Guadeloupe", "Samples"]
### reads
compute_read_counts(ps.quinque.ovary %>% subset_samples(Strain=="Field - Guadeloupe")) -> df3["Field - Guadeloupe", "Reads"]


## Wolbachia -
### oligotype
# nrow(otu_table(ps.quinque.ovary %>% 
#                  subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)") %>% 
#                  check_ps())) -> df3["Laboratory - Slab TC (Wolbachia -)", "Oligotypes"]
### samples
# nrow(ps.quinque.ovary@sam_data %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df3["Laboratory - Slab TC (Wolbachia -)", "Samples"]
# ### reads
# compute_read_counts(ps.quinque.ovary %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df3["Laboratory - Slab TC (Wolbachia -)", "Reads"]



# Aedes aegypti
ps.aedes.ovary <- subset_samples(ps.aedes, Organ=="Ovary")
ps.aedes.ovary <- check_ps(ps.aedes.ovary)

## Guadeloupe
### oligotype
nrow(otu_table(ps.aedes.ovary %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df3["Aedes aegypti (Guadeloupe)", "Oligotypes"]
### samples
nrow(ps.aedes.ovary@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df3["Aedes aegypti (Guadeloupe)", "Samples"]
### reads
compute_read_counts(ps.aedes.ovary %>% subset_samples(Strain=="Field - Guadeloupe")) -> df3["Aedes aegypti (Guadeloupe)", "Reads"]


# Total
ps.ovary <- ps.filter %>% subset_samples(Organ=="Ovary")
ps.ovary <- ps.ovary %>% check_ps()
### oligotype
nrow(ps.ovary@otu_table) -> df3["Total", "Oligotypes"]
### samples
nrow(ps.ovary@sam_data) -> df3["Total", "Samples"]
### reads
compute_read_counts(ps.ovary) -> df3["Total", "Reads"]

df3[is.na(df3)] <- 0

df3 %>% 
  kbl() %>%
  kable_paper("hover", full_width = F)
```


## Plot of number of reads by sample 
```{r fig.height = 12, fig.width=14, warning=FALSE}
sample_data(ps.filter)$Read_depth <- sample_sums(ps.filter)
metadata_read <- data.frame(ps.filter@sam_data)

p <- ggplot(metadata_read, aes(x = Sample, y = Read_depth))+ 
  geom_bar(position = "dodge", stat = "identity")+
    scale_fill_manual(values = col)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("") + 
  theme(legend.title = element_text(size = 18), 
        legend.position="bottom",
        legend.text=element_text(size=11), 
        panel.spacing.y=unit(1, "lines"), 
        panel.spacing.x=unit(0.8, "lines"),
        panel.spacing=unit(0,"lines"),
        strip.background=element_rect(color="grey30", fill="grey90"),
        panel.border=element_rect(color="grey90"),
        axis.ticks.x=element_blank(),
        strip.text.x = element_text(size = 12)) +
 facet_wrap(~Species_italic+Strain_italic+Organ, scales = "free_x", ncol=4, labeller=label_parsed)+
  labs(y="Sequence counts")+
  ylim(0, 550000)+
  geom_text(aes(label=Read_depth), position=position_dodge(width=0.5), size=2, hjust=-0.25, vjust=0.25, angle=90)+
  guides(fill=guide_legend(title="Oligotype", label.theme = element_text(size = 10, face = "italic", colour = "Black", angle = 0)))

p
```

# Filter samples 

## Remove sample with number of reads <1000
```{r}
a <- prune_samples(sample_sums(ps.filter)<=1000, ps.filter)
test <- data.frame(a@sam_data)
test <- test[test$Sample!="NP20" & test$Sample!="NP29" & test$Sample!="NP30" & test$Sample!="NP34" & test$Sample!="NP36",]
toremove <- test$Sample

ps.filter <- subset_samples(ps.filter, !(Sample %in% toremove))
ps.filter
compute_read_counts(ps.filter)
```

## Plot of final number of reads by sample
```{r fig.height = 22, fig.width=18}

sample_data(ps.filter)$Read_depth <- sample_sums(ps.filter)
metadata_read <- data.frame(ps.filter@sam_data)
metadata_read_whole <- metadata_read[metadata_read$Organ=="Whole",]
metadata_read_ovary <- metadata_read[metadata_read$Organ=="Ovary",]


guide_italics <- guides(fill = guide_legend(label.theme = element_text(size = 10, face = "italic", colour = "Black", angle = 0)))

p <- ggplot(metadata_read_whole, aes(x = Sample, y = Read_depth))+ 
  geom_bar(position = "dodge", stat = "identity")+
    scale_fill_manual(values = col)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust=1, hjust=1, size=15)) +
  ggtitle("") + 
  guide_italics+
  theme(legend.title = element_text(size = 18), 
        legend.position="bottom",
        legend.text=element_text(size=16), 
        panel.spacing.y=unit(1, "lines"), 
        panel.spacing.x=unit(0.8, "lines"),
        panel.spacing=unit(0,"lines"),
        strip.background=element_rect(color="grey30", fill="grey90"),
        panel.border=element_rect(color="grey90"),
        axis.ticks.x=element_blank(),
        strip.text.x = element_text(size = 16),
        axis.text = element_text(size = 15)) +
 facet_wrap(~Species_italic+Strain_italic+Organ, scales = "free_x", ncol=3, labeller=label_parsed)+
  labs(y="Sequence counts")+
  ylim(0, 1000000)+
  geom_text(aes(label=Read_depth), position=position_dodge(width=0.5), size=4, angle=90, hjust=-0.1, vjust=0.25)+
  guides(fill=guide_legend(title="Oligotype"))

p


p2 <- ggplot(metadata_read_ovary, aes(x = Sample, y = Read_depth))+ 
  geom_bar(position = "dodge", stat = "identity")+
    scale_fill_manual(values = col)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust=1, hjust=1, size=15)) +
  ggtitle("") + 
  guide_italics+
  theme(legend.title = element_text(size = 18), 
        legend.position="bottom",
        legend.text=element_text(size=16), 
        panel.spacing.y=unit(1, "lines"), 
        panel.spacing.x=unit(0.8, "lines"),
        panel.spacing=unit(0,"lines"),
        strip.background=element_rect(color="grey30", fill="grey90"),
        panel.border=element_rect(color="grey90"),
        axis.ticks.x=element_blank(),
        strip.text.x = element_text(size = 16),
        axis.text = element_text(size = 15)) +
 facet_wrap(~Species_italic+Strain_italic+Organ, scales = "free_x", ncol=3, labeller=label_parsed)+
  labs(y="Sequence counts")+
  ylim(0, 1000000)+
  geom_text(aes(label=Read_depth), position=position_dodge(width=0.5), size=4, angle=90, hjust=-0.1, vjust=0.25)+
  guides(fill=guide_legend(title="Oligotype"))

p2


# panels
p_group <- plot_grid(p+theme(legend.position="none"), 
          p2+theme(legend.position="none"), 
          nrow=2, 
          ncol=1)+
    draw_plot_label(c("A", "B"), c(0, 0), c(1, 0.5), size = 15)

legend_plot <- get_legend(p + theme(legend.position="bottom"))

p_counts <- plot_grid(p_group, legend_plot, nrow=2, ncol=1, rel_heights = c(1, .1))
p_counts

```

# Counts after removing samples

## Create a dataframe

```{r}
x <- c("Culex pipiens", "Field - Bosc", "Field - Camping Europe", "Laboratory - Lavar", 
       "Culex quinquefasciatus", "Field - Guadeloupe", "Laboratory - Slab TC (Wolbachia -)",
       "Aedes aegypti (Guadeloupe)",
       "Total")
y <- c("Reads", "Oligotypes", "Samples")

df <- data.frame(matrix(ncol=3, nrow=9))
rownames(df) <- x
colnames(df) <- y

df2 <- df
df3 <- df
```

## Whole + Ovary
```{r}
# Culex pipiens
ps.pipiens <- subset_samples(ps.filter, Species=="Culex pipiens")
ps.pipiens <- check_ps(ps.pipiens)

## All Strain
### oligotype
nrow(ps.pipiens@otu_table) -> df["Culex pipiens", "Oligotypes"]
### samples
nrow(ps.pipiens@sam_data) -> df["Culex pipiens", "Samples"]
### reads
compute_read_counts(ps.pipiens) -> df["Culex pipiens", "Reads"]


## Bosc
### oligotype
nrow(otu_table(ps.pipiens %>% 
                 subset_samples(Strain=="Field - Bosc") %>% 
                 check_ps())) -> df["Field - Bosc", "Oligotypes"]
### samples
nrow(ps.pipiens@sam_data %>% subset_samples(Strain=="Field - Bosc")) -> df["Field - Bosc", "Samples"]
### reads
compute_read_counts(ps.pipiens %>% subset_samples(Strain=="Field - Bosc")) -> df["Field - Bosc", "Reads"]


## Camping Europe
### oligotype
nrow(otu_table(ps.pipiens %>% 
                 subset_samples(Strain=="Field - Camping Europe") %>% 
                 check_ps())) -> df["Field - Camping Europe", "Oligotypes"]
### samples
nrow(ps.pipiens@sam_data %>% subset_samples(Strain=="Field - Camping Europe")) -> df["Field - Camping Europe", "Samples"]
### reads
compute_read_counts(ps.pipiens %>% subset_samples(Strain=="Field - Camping Europe")) -> df["Field - Camping Europe", "Reads"]

## Lavar (labo)
### oligotype
nrow(otu_table(ps.pipiens %>% 
                 subset_samples(Strain=="Laboratory - Lavar") %>% 
                 check_ps())) -> df["Laboratory - Lavar", "Oligotypes"]
### samples
nrow(ps.pipiens@sam_data %>% subset_samples(Strain=="Laboratory - Lavar")) -> df["Laboratory - Lavar", "Samples"]
### reads
compute_read_counts(ps.pipiens %>% subset_samples(Strain=="Laboratory - Lavar")) -> df["Laboratory - Lavar", "Reads"]


# Culex quinquefasciatus
ps.quinque <- subset_samples(ps.filter, Species=="Culex quinquefasciatus")
ps.quinque <- check_ps(ps.quinque)

## All Strain
### oligotype
nrow(ps.quinque@otu_table) -> df["Culex quinquefasciatus", "Oligotypes"]
### samples
nrow(ps.quinque@sam_data) -> df["Culex quinquefasciatus", "Samples"]
### reads
compute_read_counts(ps.quinque) -> df["Culex quinquefasciatus", "Reads"]


## Guadeloupe
### oligotype
nrow(otu_table(ps.quinque %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df["Field - Guadeloupe", "Oligotypes"]
### samples
nrow(ps.quinque@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df["Field - Guadeloupe", "Samples"]
### reads
compute_read_counts(ps.quinque %>% subset_samples(Strain=="Field - Guadeloupe")) -> df["Field - Guadeloupe", "Reads"]


## Wolbachia -
### oligotype
nrow(otu_table(ps.quinque %>% 
                 subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)") %>% 
                 check_ps())) -> df["Laboratory - Slab TC (Wolbachia -)", "Oligotypes"]
### samples
nrow(ps.quinque@sam_data %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df["Laboratory - Slab TC (Wolbachia -)", "Samples"]
### reads
compute_read_counts(ps.quinque %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df["Laboratory - Slab TC (Wolbachia -)", "Reads"]


# Aedes aegypti
ps.aedes <- subset_samples(ps.filter, Species=="Aedes aegypti")
ps.aedes <- check_ps(ps.aedes)

## Guadeloupe
### oligotype
nrow(otu_table(ps.aedes %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df["Aedes aegypti (Guadeloupe)", "Oligotypes"]
### samples
nrow(ps.aedes@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df["Aedes aegypti (Guadeloupe)", "Samples"]
### reads
compute_read_counts(ps.aedes %>% subset_samples(Strain=="Field - Guadeloupe")) -> df["Aedes aegypti (Guadeloupe)", "Reads"]

# Total
### oligotype
nrow(ps.filter@otu_table) -> df["Total", "Oligotypes"]
### samples
nrow(ps.filter@sam_data) -> df["Total", "Samples"]
### reads
compute_read_counts(ps.filter) -> df["Total", "Reads"]

df %>% 
  kbl() %>%
  kable_paper("hover", full_width = F)
```


## Whole
```{r}
# Culex pipiens
ps.pipiens.whole <- ps.pipiens %>% subset_samples(Organ=="Whole")
ps.pipiens.whole <- ps.pipiens.whole %>% check_ps()

## All Strain
### oligotype
nrow(ps.pipiens.whole@otu_table) -> df2["Culex pipiens", "Oligotypes"]
### samples
nrow(ps.pipiens.whole@sam_data) -> df2["Culex pipiens", "Samples"]
### reads
compute_read_counts(ps.pipiens.whole) -> df2["Culex pipiens", "Reads"]


## Bosc
### oligotype
nrow(otu_table(ps.pipiens.whole %>% 
                 subset_samples(Strain=="Field - Bosc") %>% 
                 check_ps())) -> df2["Field - Bosc", "Oligotypes"]
### samples
nrow(ps.pipiens.whole@sam_data %>% subset_samples(Strain=="Field - Bosc")) -> df2["Field - Bosc", "Samples"]
### reads
compute_read_counts(ps.pipiens.whole %>% subset_samples(Strain=="Field - Bosc")) -> df2["Field - Bosc", "Reads"]


## Camping Europe
### oligotype
nrow(otu_table(ps.pipiens.whole %>% 
                 subset_samples(Strain=="Field - Camping Europe") %>% 
                 check_ps())) -> df2["Field - Camping Europe", "Oligotypes"]
### samples
nrow(ps.pipiens.whole@sam_data %>% subset_samples(Strain=="Field - Camping Europe")) -> df2["Field - Camping Europe", "Samples"]
### reads
compute_read_counts(ps.pipiens.whole %>% subset_samples(Strain=="Field - Camping Europe")) -> df2["Field - Camping Europe", "Reads"]

## Lavar (labo)
### oligotype
nrow(otu_table(ps.pipiens.whole %>% 
                 subset_samples(Strain=="Laboratory - Lavar") %>% 
                 check_ps())) -> df2["Laboratory - Lavar", "Oligotypes"]
### samples
nrow(ps.pipiens.whole@sam_data %>% subset_samples(Strain=="Laboratory - Lavar")) -> df2["Laboratory - Lavar", "Samples"]
### reads
compute_read_counts(ps.pipiens.whole %>% subset_samples(Strain=="Laboratory - Lavar")) -> df2["Laboratory - Lavar", "Reads"]



# Culex quinquefasciatus
ps.quinque.whole <- subset_samples(ps.quinque, Organ=="Whole")
ps.quinque.whole <- check_ps(ps.quinque.whole)

## All Strain
### oligotype
nrow(ps.quinque.whole@otu_table) -> df2["Culex quinquefasciatus", "Oligotypes"]
### samples
nrow(ps.quinque.whole@sam_data) -> df2["Culex quinquefasciatus", "Samples"]
### reads
compute_read_counts(ps.quinque.whole) -> df2["Culex quinquefasciatus", "Reads"]


## Guadeloupe
### oligotype
nrow(otu_table(ps.quinque.whole %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df2["Field - Guadeloupe", "Oligotypes"]
### samples
nrow(ps.quinque.whole@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df2["Field - Guadeloupe", "Samples"]
### reads
compute_read_counts(ps.quinque.whole %>% subset_samples(Strain=="Field - Guadeloupe")) -> df2["Field - Guadeloupe", "Reads"]


## Wolbachia -
### oligotype
nrow(otu_table(ps.quinque.whole %>% 
                 subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)") %>% 
                 check_ps())) -> df2["Laboratory - Slab TC (Wolbachia -)", "Oligotypes"]
### samples
nrow(ps.quinque.whole@sam_data %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df2["Laboratory - Slab TC (Wolbachia -)", "Samples"]
### reads
compute_read_counts(ps.quinque.whole %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df2["Laboratory - Slab TC (Wolbachia -)", "Reads"]



# Aedes aegypti
ps.aedes.whole <- subset_samples(ps.aedes, Organ=="Whole")
ps.aedes.whole <- check_ps(ps.aedes.whole)

## Guadeloupe
### oligotype
nrow(otu_table(ps.aedes.whole %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df2["Aedes aegypti (Guadeloupe)", "Oligotypes"]
### samples
nrow(ps.aedes.whole@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df2["Aedes aegypti (Guadeloupe)", "Samples"]
### reads
compute_read_counts(ps.aedes.whole %>% subset_samples(Strain=="Field - Guadeloupe")) -> df2["Aedes aegypti (Guadeloupe)", "Reads"]


# Total
ps.whole <- ps.filter %>% subset_samples(Organ=="Whole")
ps.whole <- ps.whole %>% check_ps()
### oligotype
nrow(ps.whole@otu_table) -> df2["Total", "Oligotypes"]
### samples
nrow(ps.whole@sam_data) -> df2["Total", "Samples"]
### reads
compute_read_counts(ps.whole) -> df2["Total", "Reads"]

df2 %>% 
  kbl() %>%
  kable_paper("hover", full_width = F)
```




## Ovary
```{r}
# Culex pipiens
ps.pipiens.ovary <- ps.pipiens %>% subset_samples(Organ=="Ovary")
ps.pipiens.ovary <- ps.pipiens.ovary %>% check_ps()

## All Strain
### oligotype
nrow(ps.pipiens.ovary@otu_table) -> df3["Culex pipiens", "Oligotypes"]
### samples
nrow(ps.pipiens.ovary@sam_data) -> df3["Culex pipiens", "Samples"]
### reads
compute_read_counts(ps.pipiens.ovary) -> df3["Culex pipiens", "Reads"]


## Bosc
### oligotype
nrow(otu_table(ps.pipiens.ovary %>% 
                 subset_samples(Strain=="Field - Bosc") %>% 
                 check_ps())) -> df3["Field - Bosc", "Oligotypes"]
### samples
nrow(ps.pipiens.ovary@sam_data %>% subset_samples(Strain=="Field - Bosc")) -> df3["Field - Bosc", "Samples"]
### reads
compute_read_counts(ps.pipiens.ovary %>% subset_samples(Strain=="Field - Bosc")) -> df3["Field - Bosc", "Reads"]


## Camping Europe
### oligotype
nrow(otu_table(ps.pipiens.ovary %>% 
                 subset_samples(Strain=="Field - Camping Europe") %>% 
                 check_ps())) -> df3["Field - Camping Europe", "Oligotypes"]
### samples
nrow(ps.pipiens.ovary@sam_data %>% subset_samples(Strain=="Field - Camping Europe")) -> df3["Field - Camping Europe", "Samples"]
### reads
compute_read_counts(ps.pipiens.ovary %>% subset_samples(Strain=="Field - Camping Europe")) -> df3["Field - Camping Europe", "Reads"]

## Lavar (labo)
### oligotype
nrow(otu_table(ps.pipiens.ovary %>% 
                 subset_samples(Strain=="Laboratory - Lavar") %>% 
                 check_ps())) -> df3["Laboratory - Lavar", "Oligotypes"]
### samples
nrow(ps.pipiens.ovary@sam_data %>% subset_samples(Strain=="Laboratory - Lavar")) -> df3["Laboratory - Lavar", "Samples"]
### reads
compute_read_counts(ps.pipiens.ovary %>% subset_samples(Strain=="Laboratory - Lavar")) -> df3["Laboratory - Lavar", "Reads"]



# Culex quinquefasciatus
ps.quinque.ovary <- subset_samples(ps.quinque, Organ=="Ovary")
ps.quinque.ovary <- check_ps(ps.quinque.ovary)

## All Strain
### oligotype
nrow(ps.quinque.ovary@otu_table) -> df3["Culex quinquefasciatus", "Oligotypes"]
### samples
nrow(ps.quinque.ovary@sam_data) -> df3["Culex quinquefasciatus", "Samples"]
### reads
compute_read_counts(ps.quinque.ovary) -> df3["Culex quinquefasciatus", "Reads"]


## Guadeloupe
### oligotype
nrow(otu_table(ps.quinque.ovary %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df3["Field - Guadeloupe", "Oligotypes"]
### samples
nrow(ps.quinque.ovary@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df3["Field - Guadeloupe", "Samples"]
### reads
compute_read_counts(ps.quinque.ovary %>% subset_samples(Strain=="Field - Guadeloupe")) -> df3["Field - Guadeloupe", "Reads"]


## Wolbachia -
### oligotype
# nrow(otu_table(ps.quinque.ovary %>% 
#                  subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)") %>% 
#                  check_ps())) -> df3["Laboratory - Slab TC (Wolbachia -)", "Oligotypes"]
### samples
# nrow(ps.quinque.ovary@sam_data %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df3["Laboratory - Slab TC (Wolbachia -)", "Samples"]
# ### reads
# compute_read_counts(ps.quinque.ovary %>% subset_samples(Strain=="Laboratory - Slab TC (Wolbachia -)")) -> df3["Laboratory - Slab TC (Wolbachia -)", "Reads"]



# Aedes aegypti
ps.aedes.ovary <- subset_samples(ps.aedes, Organ=="Ovary")
ps.aedes.ovary <- check_ps(ps.aedes.ovary)

## Guadeloupe
### oligotype
nrow(otu_table(ps.aedes.ovary %>% 
                 subset_samples(Strain=="Field - Guadeloupe") %>% 
                 check_ps())) -> df3["Aedes aegypti (Guadeloupe)", "Oligotypes"]
### samples
nrow(ps.aedes.ovary@sam_data %>% subset_samples(Strain=="Field - Guadeloupe")) -> df3["Aedes aegypti (Guadeloupe)", "Samples"]
### reads
compute_read_counts(ps.aedes.ovary %>% subset_samples(Strain=="Field - Guadeloupe")) -> df3["Aedes aegypti (Guadeloupe)", "Reads"]


# Total
ps.ovary <- ps.filter %>% subset_samples(Organ=="Ovary")
ps.ovary <- ps.ovary %>% check_ps()
### oligotype
nrow(ps.ovary@otu_table) -> df3["Total", "Oligotypes"]
### samples
nrow(ps.ovary@sam_data) -> df3["Total", "Samples"]
### reads
compute_read_counts(ps.ovary) -> df3["Total", "Reads"]

df3[is.na(df3)] <- 0

df3 %>% 
  kbl() %>%
  kable_paper("hover", full_width = F)
```

# Save data 

## Save count tables and count plots
```{r warning=FALSE}
write.table(df, "../../../output/1_MED/1D/1D_Counts_all.tsv", sep="\t", row.names = TRUE, col.names=NA)
write.table(df2, "../../../output/1_MED/1D/1D_Counts_whole.tsv", sep="\t", row.names = TRUE, col.names=NA)
write.table(df3, "../../../output/1_MED/1D/1D_Counts_ovary.tsv", sep="\t", row.names = TRUE, col.names=NA)

tiff("../../../output/1_MED/1D/1D_counts_by_sample.tiff", units="in", width=20, height=18, res=300)
p_counts
dev.off()

png("../../../output/1_MED/1D/1D_counts_by_sample.png", units="in", width=20, height=18, res=300)
p_counts
dev.off()
```

## Save metadata (for Supplementary Table 1)
```{r warning=FALSE}
count_after_decontam <- data.frame(ps.filter@otu_table)
tax_after_decontam <- data.frame(ps.filter@tax_table)
metadata_after_decontam <- data.frame(ps.filter@sam_data)

compute_read_counts(ps.filter)

write.table(count_after_decontam, "../../../output/1_MED/1D/1D_count_supplementary_table1.tsv", sep="\t", row.names = TRUE, col.names=NA)
write.table(tax_after_decontam, "../../../output/1_MED/1D/1D_tax_supplementary_table1.tsv", sep="\t", row.names = TRUE, col.names=NA)
```

## Save phyloseq object after decontam
```{r warning=FALSE}
saveRDS(ps.filter, "../../../output/1_MED/1D/1D_MED_phyloseq_decontam.rds")
```

# Session info
```{r warning=FALSE, message=FALSE}
sessionInfo()
```

