---
title: "Filter data and run MED"
author: "Hans Schrieke and Julie Reveillaud"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    lightbox: true
    gallery: false
    fig_caption: true
    highlight: tango
    toc_depth: 3
    code_folding: "show"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Paths and libraries setting

```{r, warning=FALSE, message=FALSE}
# load main custom functions
source("../../source/functions.R")

# load packages
packages <- c("tidyverse", "kableExtra", "Biostrings", "seqinr", "ggplot2", "openxlsx")
invisible(lapply(packages, require, character.only = TRUE))

# create output folders if needed
if (!dir.exists("../../../output/1_MED")) {dir.create("../../../output/1_MED")}
if (!dir.exists("../../../output/1_MED/1A")) {dir.create("../../../output/1_MED/1A")}

# variable for path
output <- "../../../"
```

# Load metadata

```{r warning=FALSE, message=FALSE}
# load Supplementary Table 1 - Sheet 1
metadata <- read.xlsx(paste0(output, "metadata/Supplementary_Table_1.xlsx"), sheet=1)

# remove special characters from Sample ID
metadata$Sample.ID <- gsub("[^0-9A-Za-z///' ]","" ,metadata$Sample.ID,  ignore.case = TRUE)

# select useful columns for workflow
metadata <- metadata %>% select(-c(7, 10:18))

# change colnames 
colnames(metadata) <- c("Sample", "Strain", "Field", "Country", "Organ", "Species", "Run", "Control")

# set rownames
rownames(metadata) <- metadata$Sample

# save this metadata table that will be used for the other scripts
write.table(metadata, paste0(output, "metadata/metadata.csv"), sep=";", row.names = FALSE)
```

# Prepare data for MED

To be run in MED workflow, data needs to be in this form (https://merenlab.org/2012/05/11/oligotyping-pipeline-explained/#preparing-the-fasta-file) : 

```{r}
cat(
" >Sample-01_ReadX
 GTTGAAAAAGTTAGTGGTGAAATCCCAGA
 >Sample-01_ReadY
 GTTGAAAAAGTTAGTGGTGAAATCCCAGA
 >Sample-01_ReadZ
 GGTGAAAAAGTTAGTGGTGAAATCCCAGA
 >Sample-02_ReadN
 GTTGAAAAAGTTAGTGGTGAAATCCCAGA
 >Sample-02_ReadM
 GTTGAAAAAGTTAGTGGTGAAATCCCAGA")
```

## Remove sample with less than 1000 reads (with exceptions)

```{r warning=FALSE, message=FALSE}
# load the fasta file
fasta_res <- load_sequence_file(paste0(output, "output/0_merge/0D_combined/"), filename="combined_file.fasta")

read_by_sample <- fasta_res[[1]]
fasta_df <- fasta_res[[2]]

# list and remove samples with less than 1000 reads except to NP20, NP29, NP30, NP34 et NP36

colnames(read_by_sample)[1] <- "Sample"
metadata <- metadata %>% merge(read_by_sample, by="Sample")
levels(metadata$Strain)

# save number of reads by sample
write.table(metadata, paste0(output, "output/1_MED/1A/1A_MED_metadata_nreads_before_filter.csv"), sep=";", row.names = FALSE)

# remove sample with less than 1000 reads except to NP20, NP29, NP30, NP34 et NP36
nmax = 1000
bad_samples <- metadata[metadata$nread<nmax,]

toremove <- c("NP20", "NP29", "NP30", "NP34", "NP36")
bad_samples <- bad_samples[!(bad_samples$Sample %in% toremove), ]

metadata <- metadata[!metadata$Sample %in% bad_samples$Sample, ]

# save number of reads by sample after filter
write.table(metadata, paste0(output, "output/1_MED/1A/1A_MED_metadata_nreads_after_filter.csv"), sep=";", row.names = FALSE)
```

## Plot of number of reads by sample
```{r fig.height = 12, fig.width=18}

p <- ggplot(metadata, aes(x = Sample, y = nread))+ 
  geom_bar(position = "dodge", stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        legend.title = element_text(size = 18), 
        legend.position="bottom",
        legend.text=element_text(size=11), 
        panel.spacing.y=unit(1, "lines"), 
        panel.spacing.x=unit(0.8, "lines"),
        panel.spacing=unit(0,"lines"),
        strip.background=element_rect(color="grey30", fill="grey90"),
        panel.border=element_rect(color="grey90"),
        axis.ticks.x=element_blank(),
        strip.text.x = element_text(size = 12)) +
  guides(fill=guide_legend(title="Sample", label.theme = element_text(size = 10, face = "italic", 
                                                                      colour = "Black", angle = 0)))+
  facet_wrap(~Species+Strain+Organ, scales = "free_x", ncol=4)+
  labs(y="Sequence counts")+
  ylim(0, 900000)+
  geom_text(aes(label=nread), 
            position=position_dodge(width=0.5), 
            size=2, hjust=-0.25, vjust=0.25, angle=90)+
  scale_fill_manual(values = col)+
  theme_bw() +
  ggtitle("") 
  
p
```

## Total number of read and samples in the final fasta file

```{r warning=FALSE}
# Merge of fasta_df and metadata
fasta_df$Sample_only <- sub("_.*", "", fasta_df$Sample)
fasta_df <- droplevels(fasta_df)

fasta_df_final <- fasta_df[fasta_df$Sample_only %in% metadata$Sample,]
fasta_df_final <- droplevels(fasta_df_final)

cat(paste0("Number of reads: ", nrow(fasta_df_final), "\n"))
cat(paste0("Number of samples: ", levels(fasta_df_final$Sample_only %>% as.factor()) %>% length(), "\n"))
```

## Save fasta and list of the remove samples
```{r warning=FALSE, message=FALSE}
# objet R
save(fasta_df_final,  file=paste0(output, "output/1_MED/1A/1A_MED_sequences.RData"))

# write fasta
write.fasta(sequences=as.list(fasta_df_final$Seq), names=fasta_df_final$Sample, file.out=paste0(output, "output/1_MED/1A/1A_MED_sequences.fasta"))

# write list of the removed samples
write.table(bad_samples, file=paste0(output, "output/1_MED/1A/1A_MED_removed_samples.tsv"), sep="\t", row.names=FALSE, quote = FALSE)
```

# Run MED

To run MED, we used a docker container where oligotyping v2.1 were installed. To install this container, you can follow this great tutorial : https://merenlab.org/2014/09/02/virtualbox/

```{r engine='bash', attr.output='style="max-height: 300px;"'}
# variable for path
OUTPUT="../../../output"
path_data="$OUTPUT/1_MED/1A"

# move to work directory
cd $path_data

# run docker session
docker run -d --name oligotyping --rm -v `pwd`:`pwd` -w `pwd` -it meren/oligotyping:2.1

# check files 
docker exec oligotyping ls 

# get info from fasta 
docker exec oligotyping o-get-sample-info-from-fasta 1A_MED_sequences.fasta

# fill gaps
docker exec oligotyping o-pad-with-gaps 1A_MED_sequences.fasta 

# get info from fasta to check if all is good after the fill gaps step
docker exec oligotyping o-get-sample-info-from-fasta 1A_MED_sequences.fasta-PADDED-WITH-GAPS

# run MED
docker exec oligotyping decompose 1A_MED_sequences.fasta-PADDED-WITH-GAPS

# kill docker container
docker container kill oligotyping
```

# Session info
```{r warning=FALSE, message=FALSE}
sessionInfo()
```


